### 2014-09-13 notes.Rmd
map seq

# Summer Data Challenge

## Intro

The challenge is an open-ended contest to find new insights using our exclusively prepared datasets, and propose how these new insights can be used to create social or economic value in the form of new products, services, or businesses.

Submissions will be judged on technical ability, creativity, practicality, and their potential for generating social or economic impact. Anybody can download our datasets and enter the contest and submissions will be judged by a panel made up of experts from academia and industry. The best submissions will be showcased on Imperial College's Data Science Institute website.

## London house prices datasets

1. Map of house sales over 5 year period. Solar theme, white flashes for house sale, brightness / persistence proportional to cost. js plotting library. Start off with ggplot2 svg then animate via d3? (this is icing on the cake really, not relevant for new insights)

2. Actionable insights. Relate percentages changes to other variables, as supplied by other datasets from e.g. guardian data. Machine learning, variable importance etc. What makes for a good local housing market?

### 2014-09-13 R/houses.R

@@ -66,12 +66,55 @@ houses$month <- months(houses$Trdate)
 houses$yearmon <- as.factor(as.yearmon(houses$Trdate))
 
 h3 <- houses %>% group_by(area, district, yearmon) %>%
-    filter(n() > 1e2) %>%
+  # only more than 30 per month
+  filter(n() > 30) %>%
   summarise(median=median(Price))
 
 ## monthly hi - low - max like open close of stocks?
 h3$yearmon <- as.Date(as.character(paste("1", h3$yearmon)), format="%d %B %Y")
 
-ggplot(h3, aes(x=yearmon, y=median, col=area)) +
+ggplot(h3, aes(x=yearmon, y=median, col=area, group=district)) +
   facet_wrap(~area, scales="free_y") +
-  geom_line()
+  geom_line() 
+
+# postcode data
+
+# match pcd with (oseast1m, osnrth1m), create smaller subset for easier usage
+#ons <- read.csv("ONSPD_MAY_2014_csv/Data/ONSPD_MAY_2014_UK.csv", stringsAsFactors=F)
+#matches <- ons[ons$pcd %in% houses$Postcode,]
+#rm(ons)
+#saveRDS(matches, "rds/ons_postcode_subset.rds")
+matches <- readRDS("rds/ons_postcode_subset.rds")
+
+m <- merge(matches[, c("pcd", "oseast1m", "osnrth1m")], houses,
+      by.x="pcd", by.y="Postcode")
+
+str(m)
+
+ggplot(m, aes(x=oseast1m, y=osnrth1m, col=log10(Price))) +
+  geom_point()
+
+library("maptools")
+library("maps")
+library("mapdata")
+library("ggplot2")
+gpclibPermit()
+places <- fortify(
+  readShapePoly("england-latest.shp/places.shp")
+  )
+
+xmin <- min(m$oseast1m)
+xmax <- max(m$oseast1m)
+map("worldHires","UK", col="gray90", fill=TRUE, xlim=c(xmin, xmax))
+
+## http://spatial.ly/2012/02/great-maps-ggplot2/
+#This step removes the axes labels etc when called in the plot.
+xquiet<- scale_x_continuous("", breaks=NA)
+yquiet<-scale_y_continuous("", breaks=NA)
+quiet<-list(xquiet, yquiet)
+
+#create base plot
+plon1<- ggplot(m, aes(x=oseast1m, y=osnrth1m))
+#ready the plot layers
[…]

### 2014-09-13 R/postcode_map.R

@@ -1,0 +1,87 @@
+library("gpclib")
+library("ggplot2")
+library("mapdata")
+library("maps")
+library("maptools")
+library("rgeos")
+
+houses <- read.csv("houseprices/london2009-2014-house-prices//Houseprice_2009_100km_London.csv", stringsAsFactors=F)
+post2lat <- read.csv("postcodes.csv")
+
+houses$postcodenospace <- gsub(" ", "", houses$Postcode)
+
+p <- merge(post2lat[,-1], houses, by.x="postcode", by.y="postcodenospace")
+
+lat.min <- min(p$latitude)
+lat.max <- max(p$latitude)
+long.min <- min(p$longitude)
+long.max <- max(p$longitude)
+
+map("worldHires", ".", col="gray90", fill=TRUE, 
+    xlim=c(long.min, long.max), ylim=c(lat.min, lat.max))
+points(p$longitude, p$latitude, pch=".")
+
+places <- readShapePoly("england-latest.shp/places.shp")
+as.data.frame(places)
+
+library("rgdal")
+places <- readOGR(dsn = "england-latest.shp", "waterways") 
+pp <- fortify(places)
+
+
+library("ggmap")
+map <- get_map(location="London", zoom=8, maptype="toner", color="bw", source="stamen")
+p1 <- ggmap(map, extent="device", darken=.6)
+p1 + geom_point(data=p[sample(1:nrow(p), 1e5),], 
+                aes(x=longitude, y=latitude, alpha=log10(Price)), col="white") +
+  scale_alpha_continuous(range=c(.01, .2)) +
+  labs(x="", y="") + theme(legend.position="none") 
+
+# filename ordering for (e.g.) imagemagick (yes, 51 not 52)
+getfn <- function(n)
+  ifelse(n > 51, paste0("zz", letters[n %% 26]), 
+         ifelse(n > 26, paste0("z", letters[n %% 26]), letters[n]))
+
+
+for( ym in 1:65 ){
+  # x=1, y=50.5
+  p1 + geom_point(data=p[p$yearmon == levels(p$yearmon)[ym],], 
+                       aes(x=longitude, y=latitude, alpha=log10(Price)), 
+                  col="white", size=.5) + 
+    labs(x="", y="") + 
+    theme(legend.position="none",
+          axis.ticks = element_blank(), axis.text.x = element_blank(),
+         axis.text.y = element_blank()) +
+    annotate("text", x=1, y=50.5, label=levels(p$yearmon)[ym], 
[…]

### 2014-09-13 R/postcode_map.R
v0 gifs

@@ -46,22 +46,22 @@ getfn <- function(n)
 for( ym in 1:65 ){
   # x=1, y=50.5
   p1 + geom_point(data=p[p$yearmon == levels(p$yearmon)[ym],], 
-                       aes(x=longitude, y=latitude, alpha=log10(Price)), 
-                  col="white", size=.5) + 
+                       aes(x=longitude, y=latitude, alpha=log10(log10(Price))), 
+                  col="white", size=.3) + 
     labs(x="", y="") + 
     theme(legend.position="none",
           axis.ticks = element_blank(), axis.text.x = element_blank(),
          axis.text.y = element_blank()) +
-    annotate("text", x=1, y=50.5, label=levels(p$yearmon)[ym], 
-             col="white", size=10, family="mono") +
-    scale_alpha_continuous(range=c(.01, .15))
+    annotate("text", x=.8, y=50.5, label=levels(p$yearmon)[ym], 
+             col="white", size=8, family="mono") +
+    scale_alpha_continuous(range=c(.2, .45))
     ggsave(paste0("plots/seq/", getfn(ym), ".png"), width=3.5, height=3.5)
 }
 
 ## stitch pngs to gif w/ imagemagick
 ## http://ubuntuforums.org/showthread.php?t=1132058
-## convert -delay 10 -loop 0 *.png -resize 75% animated.gif
-## or something more exotic, median filtering:
+## convert -delay 13 -loop 0 *.png -resize 60% animated2.gif
+## or something more exotic, median filtering, etc.:
 ## convert -delay 10 -loop 0 *.png -median 5x5 animated3.gif
 
 ## Zoom to london only
@@ -69,16 +69,14 @@ map2 <- get_map(location="London", zoom=10, maptype="toner", color="bw", source=
 p2 <- ggmap(map2, extent="device", darken=.6)
 
 for( ym in 1:65 ){
-  ym = 1
   p2 + geom_point(data=p[p$yearmon == levels(p$yearmon)[ym],], 
                   aes(x=longitude, y=latitude, alpha=log10(Price)), 
                   col="white", size=.5) + 
     labs(x="", y="") + 
     theme(legend.position="none",
           axis.ticks = element_blank(), axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          plot.title=element_text(family="mono", size=14)) +
-  annotate("text", x=.1, y=51.2, label=levels(p$yearmon)[ym], col="white") +
[…]

### 2014-09-14 notes.Rmd
app engine, writeup website

@@
 What makes for a good local housing market?
+
+## New product OR new service OR new business
+
+Housing sale data, of interest to REITs, realtors, consumer housing market, individuals.
+
+Product ideas:
+
+* app for personal use.
+History of house sales in a given postcode, outlook for future growth, summary stats from other sources.
+Who'd use this?
+Eventually be bought out by Zoopla etc.
+for integration with their services.
+Do they offer this stuff already?
+  * Currently rightmove lists past house sales for an area
+  * ONS offers clunky postcode datasets: http://www.neighbourhood.statistics.gov.uk/dissemination/LeadAreaSearch.do?a=7&i=1001&m=0&s=1410700363772&enc=1&areaSearchText=nw9+6ef&extendedList=false&searchAreas=
+  * Crime statistics: http://www.crime-statistics.co.uk/postcode
+
+Input postcode of interest, get nice visualisation of previous house sales, median rolling prices. Integrated with other data. Also offer surrouding postcode data for comparison.

### 2014-09-14 R/datablog.R

@@ -1,0 +1,27 @@
+boroughs <- read.csv("datablog_data/London boroughs in numbers.csv")
+
+# relate boroughs to postcodes via london_postcodes.csv from: http://www.doogal.co.uk/london_postcodes.php
+b2pc <- read.csv("London_postcodes.csv")
+head(b2pc)
+boroughs$BOROUGH
+
+# b2pc$District \approx boroughs
+
+for( b in unique(boroughs$BOROUGH)) {
+  cat(b, "\n")
+  m <- unique(b2pc[agrep(b, b2pc$District),]$District)
+  cat(m, b, "\n") 
+}
+
+b2pc$District[agrep(boroughs$BOROUGH[1], b2pc$District)]
+
+
+earnings <- read.csv("datablog_data/Figure 3.15, Trends in median full-time gross weekly earnings of employees by gender, 1966-2008.xls.csv")
+
+london.wards <- readShapePoly("~/TD/london_wards2013/london_wards2013.shp"
+                              , proj4string=CRS(projString))
+wards.count <- nrow(london.wards@data)
+# assign id for each lsoa
+
+london.wards@data$id <- 1:wards.count
+wards.fort <- fortify(london.wards, region='id')

### 2014-09-14 R/functions.R

@@ -1,0 +1,33 @@
+## Repo of functions for use across scripts
+library("gpclib")
+library("ggplot2")
+library("mapdata")
+library("maps")
+library("maptools")
+library("rgeos")
+library("zoo")
+
+loadHouses <- function(){
+  h <- read.csv("houseprices/london2009-2014-house-prices//Houseprice_2009_100km_London.csv", 
+                stringsAsFactors=F)
+  
+  # remove underscores, convert to number
+  h$Price <- as.numeric(gsub("_", "", h$Price))
+  h$Trdate <- as.Date(h$Trdate)
+  
+  # area level (e.g. NW)
+  h$area <- factor(gsub("^(\\D+?)\\d.*", "\\1", as.character(h$Postcode)))
+  
+  # district (e.g. NW9)
+  h$district <- factor(do.call(rbind, strsplit(h$Postcode, " "))[,1])
+  
+  # sector (e.g. NW9 6)
+  h$sector <- paste(as.character(h$district),
+                    gsub(".*?\\s(\\d+).+", "\\1", h$Postcode), sep=" ")
+  
+  h$month <- months(h$Trdate)
+  h$yearmon <- as.factor(as.yearmon(h$Trdate))
+  
+  return(h)
+}
+

### 2014-09-14 R/houses.R

@@ -5,26 +5,9 @@ library("dplyr")
 library("ggplot2")
 #devtools::install_github("blmR", "blmoore")
 library("blmR")
+source("R/functions.R")
 
-houses <- read.csv("houseprices/london2009-2014-house-prices//Houseprice_2009_100km_London.csv", stringsAsFactors=F)
-
-houses$Price <- as.numeric(gsub("_", "", houses$Price))
-houses$Trdate <- as.Date(houses$Trdate)
-
-# how have house prices changed over time
-#ggplot(houses[sample(1:nrow(houses), 1e5),], 
-#       aes(x=Trdate, y=Price, col=Property_Type)) +
-#  geom_point() + scale_y_log10()
-
-# area level (e.g. NW)
-houses$area <- factor(gsub("^(\\D+?)\\d.*", "\\1", as.character(houses$Postcode)))
-
-# district (e.g. NW9)
-houses$district <- factor(do.call(rbind, strsplit(houses$Postcode, " "))[,1])
-
-# sector (e.g. NW9 6)
-houses$sector <- paste(as.character(houses$district),
-  gsub(".*?\\s(\\d+).+", "\\1", houses$Postcode), sep=" ")
+houses <- loadHouses()
 
 h2 <- group_by(houses, district) %>%
   filter(n() > 5e3)

### 2014-09-14 R/postcode_map.R

@@ -1,39 +1,30 @@
-library("gpclib")
-library("ggplot2")
-library("mapdata")
-library("maps")
-library("maptools")
-library("rgeos")
+source("R/functions.R")
 
-houses <- read.csv("houseprices/london2009-2014-house-prices//Houseprice_2009_100km_London.csv", stringsAsFactors=F)
-post2lat <- read.csv("postcodes.csv")
+houses <- loadHouses()
 
-houses$postcodenospace <- gsub(" ", "", houses$Postcode)
+lat.min <- min(houses$Latitude)
+lat.max <- max(houses$Latitude)
+long.min <- min(houses$Longitude)
+long.max <- max(houses$Longitude)
 
-p <- merge(post2lat[,-1], houses, by.x="postcode", by.y="postcodenospace")
-
-lat.min <- min(p$latitude)
-lat.max <- max(p$latitude)
-long.min <- min(p$longitude)
-long.max <- max(p$longitude)
-
+# base graphics data overview
 map("worldHires", ".", col="gray90", fill=TRUE, 
     xlim=c(long.min, long.max), ylim=c(lat.min, lat.max))
 points(p$longitude, p$latitude, pch=".")
 
-places <- readShapePoly("england-latest.shp/places.shp")
-as.data.frame(places)
-
-library("rgdal")
-places <- readOGR(dsn = "england-latest.shp", "waterways") 
-pp <- fortify(places)
+# places <- readShapePoly("england-latest.shp/places.shp")
+# as.data.frame(places)
+# 
+# library("rgdal")
+# places <- readOGR(dsn = "england-latest.shp", "waterways") 
+# pp <- fortify(places)
 
 
 library("ggmap")
 map <- get_map(location="London", zoom=8, maptype="toner", color="bw", source="stamen")
 p1 <- ggmap(map, extent="device", darken=.6)
-p1 + geom_point(data=p[sample(1:nrow(p), 1e5),], 
-                aes(x=longitude, y=latitude, alpha=log10(Price)), col="white") +
+p1 + geom_point(data=houses[sample(1:nrow(houses), 1e5),], 
+                aes(x=Longitude, y=Latitude, alpha=log10(Price)), col="white") +
   scale_alpha_continuous(range=c(.01, .2)) +
   labs(x="", y="") + theme(legend.position="none") 
 
@@ -45,14 +36,14 @@ getfn <- function(n)
 
 for( ym in 1:65 ){
   # x=1, y=50.5
[…]

### 2014-09-14 app_cmds.sh

@@ -1,0 +1,5 @@
+## run locally at http://localhost:8080
+dev_appserver.py appengine
+
+## deploy to model-nexus-703.appspot.com
+appcfg.py -A model-nexus-703 update appengine

### 2014-09-14 appengine/hello/views.py

@@ -1,0 +1,4 @@
+from django import http
+
+def home(request):
+    return http.HttpResponse('Hello World!')

### 2014-09-14 writeup/js/init.js

@@ -1,0 +1,126 @@
+/*
+	Prologue by HTML5 UP
+	html5up.net | @n33co
+	Free for personal and commercial use under the CCA 3.0 license (html5up.net/license)
+*/
+
+(function($) {
+
+	skel.init({
+		reset: 'full',
+		breakpoints: {
+			'global':	{ range: '*', href: 'css/style.css', containers: 1400, grid: { gutters: 40 }, viewport: { scalable: false } },
+			'wide':		{ range: '961-1880', href: 'css/style-wide.css', containers: 1200, grid: { gutters: 40 } },
+			'normal':	{ range: '961-1620', href: 'css/style-normal.css', containers: 960, grid: { gutters: 40 } },
+			'narrow':	{ range: '961-1320', href: 'css/style-narrow.css', containers: '100%', grid: { gutters: 20 } },
+			'narrower':	{ range: '-960', href: 'css/style-narrower.css', containers: '100%', grid: { gutters: 15 } },
+			'mobile':	{ range: '-640', href: 'css/style-mobile.css', grid: { collapse: true } }
+		}
+	}, {
+		layers: {
+			layers: {
+				sidePanel: {
+					hidden: true,
+					breakpoints: 'narrower',
+					position: 'top-left',
+					side: 'left',
+					animation: 'pushX',
+					width: 240,
+					height: '100%',
+					clickToClose: true,
+					html: '<div data-action="moveElement" data-args="header"></div>',
+					orientation: 'vertical'
+				},
+				sidePanelToggle: {
+					breakpoints: 'narrower',
+					position: 'top-left',
+					side: 'top',
+					height: '4em',
+					width: '5em',
+					html: '<div data-action="toggleLayer" data-args="sidePanel" class="toggle"></div>'
+				}
+			}
+		}
+	});
+
+	$(function() {
+
+		var	$window = $(window),
+			$body = $('body');
+			
+		// Disable animations/transitions until the page has loaded.
+			$body.addClass('is-loading');
+			
+			$window.on('load', function() {
+				$body.removeClass('is-loading');
+			});
+			
+		// CSS polyfills (IE<9).
+			if (skel.vars.IEVersion < 9)
+				$(':last-child').addClass('last-child');
+
+		// Forms (IE<10).
+			var $form = $('form');
+			if ($form.length > 0) {
+
+				$form.find('.form-button-submit')
+					.on('click', function() {
+						$(this).parents('form').submit();
+						return false;
+					});
+
+				if (skel.vars.IEVersion < 10) {
[…]

### 2014-09-28 R/functions.R
graph 1, google map addition

@@ -90,7 +90,7 @@ theme_sdc_null <- function(base_size=12, base_family="sans") {
     panel.grid.minor.y = element_line(size=.1),
     ## transparent bg?
     plot.background = element_rect(fill = "transparent", colour = NA),
-    plot.title = element_text(color="#ffa775", face=2, size=16),
+    plot.title = element_text(color="grey60", face=2, size=20),
     plot.margin = grid::unit(c(1, 1, 0.5, 0.5), "lines"),
     strip.background = element_rect(fill="grey90", color="grey90"),
     strip.text = element_text(color="grey25"),

### 2014-09-28 R/postcode_map.R

@@ -12,12 +12,6 @@ map("worldHires", ".", col="gray90", fill=TRUE,
     xlim=c(long.min, long.max), ylim=c(lat.min, lat.max))
 points(p$longitude, p$latitude, pch=".")
 
-# places <- readShapePoly("england-latest.shp/places.shp")
-# as.data.frame(places)
-# 
-# library("rgdal")
-# places <- readOGR(dsn = "england-latest.shp", "waterways") 
-# pp <- fortify(places)
 
 
 library("ggmap")

### 2014-09-28 R/single_postcode.R

@@ -1,5 +1,8 @@
 library("dplyr")
 library("ggplot2")
+library("RColorBrewer")
+library("reshape2")
+source("R/functions.R")
 options(scipen=99)
 
 ## Cases per regions       Cases  Median n
@@ -14,29 +17,10 @@ options(scipen=99)
 top <- sort(table(houses$district), decreasing=T)
 head(top)
 
+# say you're looking at a house in: SW18 4HU 
 
 sw18 <- subset(houses, district == "SW18")
-ggplot(sw18, aes(x=Trdate, y=Price, col=Newbuild)) +
-  facet_wrap(~Newbuild) +
-  #geom_point() + 
-  geom_smooth(method="lm")
-
-
-counts <- group_by(houses, Postcode) %>% summarise(count=n())
-ggplot(counts, aes(x=count)) + 
-  stat_ecdf(geom="line") +
-  scale_x_log10()
-
-dcounts <- group_by(houses, district) %>% summarise(count=n())
-ggplot(dcounts, aes(x=count)) + 
-  stat_ecdf(geom="line") +
-  scale_x_log10()
-
-ggplot(houses, aes(x=Newbuild, y=Trdate)) +
-  geom_violin()
-  
-ggplot(houses, aes(x=Trdate, y=Price)) +
-  facet_wrap(~Newbuild)
+sw184 <- subset(houses, sector == "SW18 4")
 
 # polar co-ord chart, month on month
 library("zoo")
@@ -49,14 +33,89 @@ ggplot(sw18, aes(x=Trdate, y=Price)) +
   #scale_y_log10() +
   geom_smooth(method="loess") +
   coord_cartesian(ylim=c(quantile(sw18$Price, .05), 
-                             quantile(sw18$Price, .95)))
+                         quantile(sw18$Price, .95)))
 
-smod <- loess(Price ~ as.numeric(Trdate), data=sw18)
-months <- seq(min(sw18$Trdate), max(sw18$Trdate), by="quarter")
-cat(paste0(signif(predict(smod, as.numeric(months)), 3), sep=","))
 
-amod <- loess(Price ~ as.numeric(Trdate), 
-              data=houses[sample(1:nrow(houses), 1e5),])
-cat(paste0(signif(predict(amod, as.numeric(months)), 3), sep=","))
+modelArea <- function(data, dates, label){
+  if (nrow(data) > 2e4){
+    out <- loess(Price ~ as.numeric(Trdate), 
[…]

### 2014-09-28 R/single_postcode.R
is fig

@@ -107,7 +107,8 @@ ggplot(pcsummary, aes(y=med/1e3, x=sales,
   scale_size_continuous(range=c(2, 5)) +
   labs(x="5yr house sales", y="Median sale price (000s £)")
 
-
+## SW18 4**
+pdf("plots/sw184_priceDist.pdf", 4, 2)
 ggplot(pcsummary, aes(x=med/1e3)) +
   geom_density(fill=I(rgb(239, 245, 249, max=255)),
                col=I(rgb(175,205,226, max=255))) + theme_sdc() +
@@ -116,6 +117,55 @@ ggplot(pcsummary, aes(x=med/1e3)) +
   theme(axis.ticks = element_blank(), 
         axis.title.y = element_blank(), 
         axis.text.y =  element_blank()) +
-  labs(x="5yr house sales (median price, 000s £)")
+  labs(x="5yr house sales (median price, 000s £)") +
+  annotate("text", x=750, y=4e-3, label="SW18 4", col=I("grey30"), size=6)
+dev.off()
 
+## SW18
+sw18.s <- group_by(sw18, sector) %>%
+  summarise(med=median(Price))
 
+sw18.s <- sw18.s[order(sw18.s$med),]
+# order by median sale
+sw18$sector <- factor(sw18$sector, levels=sw18.s$sector)
+
+pdf("plots/sw18_priceDist.pdf", 5, 3.5)
+ggplot(sw18, aes(x=sector, y=Price/1e3)) +
+  geom_violin(fill=I(rgb(239, 245, 249, max=255)),
+              col=I(rgb(175,205,226, max=255))) + 
+                geom_jitter(alpha=I(.35), col=I("grey80"), size=I(.8)) +
+  theme_sdc() + scale_y_log10(breaks=c(100,250,500,1000,2500,5000)) +
+  labs(y="House sales (000s £)", x="") +
+  annotate("text", x="SW18 4", y=7.5e3, label="SW18", col=I("grey40"), size=6)
+dev.off()
+
+london.s
+
+## London 
+london.s <- group_by(ldn, area) %>% 
+  summarise(med=median(Price))
+london.s <- london.s[order(london.s$med),]
+ldn$area <- factor(ldn$area, levels=(london.s$area))
+
+pdf("plots/ldn_areas.pdf", 3.5, 5)
+ggplot(ldn, aes(x=area, y=Price/1e3)) +
+  geom_violin(fill=I(rgb(203, 212, 231, max=255)),
[…]

### 2014-09-29 R/clusterPrices.R
tweaks

@@ -1,0 +1,18 @@
+source("R/functions.R")
+set.seed(2)
+
+## cluster monthly price change profiles
+houses <- loadHouses()
+
+
+# sample 20 random postcodes for clustering
+# redraw london as clustered by price
+pclist <- unique(houses$sector)
+pclist <- pclist[sample(1:length(pclist), 20)]
+
+sub <- houses[houses$sector %in% pclist,]
+
+ggplot(sub, aes(x=yearmon, y=Price/1e3, group=sector)) +
+  geom_smooth(method="loess") + geom_point() +
+  facet_wrap(~sector, ncol=5, scales="free_y") +
+  theme_sdc()

### 2014-09-29 R/functions.R

@@ -141,3 +141,12 @@ theme_sdc <- function(base_size=12, base_family="", use_sizes=TRUE) {
 
 ## End ggplot2 theme
 
+## blank theme, from: https://gist.github.com/dsparks/3711166#file-new_theme_empty-r
+new_theme_empty <- theme_bw()
+new_theme_empty$line <- element_blank()
+new_theme_empty$rect <- element_blank()
+new_theme_empty$strip.text <- element_blank()
+new_theme_empty$axis.text <- element_blank()
+new_theme_empty$plot.title <- element_blank()
+new_theme_empty$axis.title <- element_blank()
+new_theme_empty$plot.margin <- structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit")

### 2014-09-29 R/single_postcode.R

@@ -1,7 +1,10 @@
+library("directlabels")
 library("dplyr")
 library("ggplot2")
 library("RColorBrewer")
 library("reshape2")
+library("rgdal")
+library("rgeos")
 source("R/functions.R")
 options(scipen=99)
 
@@ -18,14 +21,8 @@ top <- sort(table(houses$district), decreasing=T)
 head(top)
 
 # say you're looking at a house in: SW18 4HU 
-
 sw18 <- subset(houses, district == "SW18")
 sw184 <- subset(houses, sector == "SW18 4")
-
-# polar co-ord chart, month on month
-library("zoo")
-z <- zoo(houses$Price, houses$yearmon)
-houses$rollmed <- rollmedian(z, 3, fill=NA)
 
 sw18 <- subset(houses, district == "SW18")
 ggplot(sw18, aes(x=Trdate, y=Price)) +
@@ -64,14 +61,10 @@ all.res <- modelArea(houses, quarts, "South-East")
 
 qs <- rbind(london.res, sw.res, sw18.res, sw184.res, sw184hu.res, all.res)
 
-#qs <- data.frame(time=quarts, sw18=sw18.q, sw=sw.q, london=l.q, all=all.q)
-#qs <- melt(qs, id.vars="time")
-
 cols <- rev(brewer.pal(5, "Blues")[-1])
 # last two: grey shaded
 cols <- c(cols, "grey60", "grey80")
 
-library("directlabels")
 qs$area <- factor(qs$area, 
             levels=c("SW18 4HU", "SW18 4", "SW18", "SW", "London", "South-East"))
 
@@ -91,7 +84,6 @@ ggplot(qs, aes(x=time, y=fit/1000, col=area, fill=area)) +
   annotate("text", x=as.Date("2010-02-01"), y=780, 
            label="SW18 4HU", size=7, col=I("grey50"))
 dev.off()
-
 
 ## compare with other postcodes in SW18 4:
 ggplot(sw184, aes(x=Trdate, y=Price, col=Postcode)) +
@@ -139,8 +131,6 @@ ggplot(sw18, aes(x=sector, y=Price/1e3)) +
   annotate("text", x="SW18 4", y=7.5e3, label="SW18", col=I("grey40"), size=6)
 dev.off()
 
-london.s
-
 ## London 
 london.s <- group_by(ldn, area) %>% 
   summarise(med=median(Price))
[…]

### 2014-09-29 app_cmds.sh
rm django stuff

@@ -1,5 +1,0 @@
-## run locally at http://localhost:8080
-dev_appserver.py appengine
-
-## deploy to model-nexus-703.appspot.com
-appcfg.py -A model-nexus-703 update appengine

### 2014-09-29 appengine/hello/models.py

@@ -1,3 +1,0 @@
-from google.appengine.ext import ndb
-
-# Create your models here.

### 2014-09-29 appengine/hello/tests.py

@@ -1,13 +1,0 @@
-# Run tests with "manage.py test".
-
-from django import http
-from django.test import TestCase
-
-from . import views
-
-class HomeViewTest(TestCase):
-    def test_home(self):
-        requst = http.HttpRequest()
-        response = views.home(request)
-        self.assertEqual(200, response.status_code)
-        self.assertIn('Hello World!', response.content)

### 2014-09-29 appengine/hello/views.py

@@ -1,4 +1,0 @@
-from django import http
-
-def home(request):
-    return http.HttpResponse('Hello World!')

### 2014-09-29 appengine/myproject/settings.py

@@ -1,147 +1,0 @@
-# Django settings for myproject project.
-
-DEBUG = True
-TEMPLATE_DEBUG = DEBUG
-
-ADMINS = (
-    # ('Your Name', 'your_email@example.com'),
-)
-
-MANAGERS = ADMINS
-
-DATABASES = {
-    'default': {
-        'ENGINE': 'django.db.backends.', # Add 'postgresql_psycopg2', 'mysql', 'sqlite3' or 'oracle'.
-        'NAME': '',                      # Or path to database file if using sqlite3.
-        # The following settings are not used with sqlite3:
-        'USER': '',
-        'PASSWORD': '',
-        'HOST': '',                      # Empty for localhost through domain sockets or '127.0.0.1' for localhost through TCP.
-        'PORT': '',                      # Set to empty string for default.
-    }
-}
-
-# Hosts/domain names that are valid for this site; required if DEBUG is False
-# See https://docs.djangoproject.com/en/1.5/ref/settings/#allowed-hosts
-ALLOWED_HOSTS = []
-
-# Local time zone for this installation. Choices can be found here:
-# http://en.wikipedia.org/wiki/List_of_tz_zones_by_name
-# although not all choices may be available on all operating systems.
-# In a Windows environment this must be set to your system time zone.
-TIME_ZONE = 'America/Chicago'
-
-# Language code for this installation. All choices can be found here:
-# http://www.i18nguy.com/unicode/language-identifiers.html
-LANGUAGE_CODE = 'en-us'
-
-SITE_ID = 1
-
-# If you set this to False, Django will make some optimizations so as not
-# to load the internationalization machinery.
-USE_I18N = True
-
-# If you set this to False, Django will not format dates, numbers and
-# calendars according to the current locale.
-USE_L10N = True
-
-# If you set this to False, Django will not use timezone-aware datetimes.
-USE_TZ = True
-
-# Absolute filesystem path to the directory that will hold user-uploaded files.
-# Example: "/var/www/example.com/media/"
-MEDIA_ROOT = ''
-
-# URL that handles the media served from MEDIA_ROOT. Make sure to use a
-# trailing slash.
-# Examples: "http://example.com/media/", "http://media.example.com/"
-MEDIA_URL = ''
-
-# Absolute path to the directory static files should be collected to.
-# Don't put anything in this directory yourself; store your static files
[…]

### 2014-09-29 appengine/myproject/urls.py

@@ -1,5 +1,0 @@
-from django.conf.urls import patterns, include, url
-
-urlpatterns = patterns('',
-    url(r'^$', 'hello.views.home'),
-)

### 2014-09-30 R/clusterPrices.R
start clustering work

@@ -16,3 +16,26 @@ ggplot(sub, aes(x=yearmon, y=Price/1e3, group=sector)) +
   geom_smooth(method="loess") + geom_point() +
   facet_wrap(~sector, ncol=5, scales="free_y") +
   theme_sdc()
+
+
+head(houses)
+library("reshape2")
+hm <- dcast(houses, sector ~ yearmon, 
+            fun.aggregate=median, value.var="Price")
+
+# length reference, mask out low cells
+hm.mask <- dcast(houses, sector ~ yearmon,  value.var="Price")
+
+image(log10(as.matrix(hm[,-1])+1))
+
+library("gplots")
+library("RColorBrewer")
+
+row.lab <- rep("", ncol(hm)-1)
+row.lab[seq(1, ncol(hm)-1, by=4)] <- colnames(hm)[-1][seq(1, ncol(hm)-1, by=4)]
+bl <- colorRampPalette(brewer.pal(9, "Blues"))(64)
+# uneven breaks, norm distributed prices
+heatmap.2(log10(as.matrix(hm[complete.cases(hm),-1])), 
+          dendrogram="row", trace="none", Colv=NULL,
+          col=bl, labRow=F, labCol=row.lab, srtCol=45)
+

### 2014-09-30 R/arima.R
ARIMA modelling, new figure

@@ -1,0 +1,86 @@
+## ARIMA forecasting
+library("dplyr")
+library("forecast")
+library("ggplot2")
+library("zoo")
+
+s <- sw184[sw184$Postcode == "SW18 4HU",]
+s <- group_by(s, Month) %>% summarise(median=median(Price))
+
+# n.b. non-stationary time series
+
+# check for (partial) auto-correlation, look at detrended etc.
+plot(swpc, lag(swpc, 12))
+
+# no detectable periodicity in monthly data
+decompose(swpc)
+seasadj(stl(swpc))
+
+# residuals OK; pacf, acf PASS
+tsdisplay(diff(swpc))
+
+# positive reasonably linear trend, can be detrended
+m <- lm(coredata(swpc) ~ index(swpc))
+detr <- zoo(resid(m), index(swpc))
+plot(swpc)
+plot(detr)
+decompose(detr)
+
+## ARIMA model fitting and forecasting
+swpc <- ts(s$median, as.numeric(as.yearmon(s$Month)))
+
+aa <- auto.arima(swpc)
+plot(swpc)
+lines(fitted(aa), col="blue")
+plot(forecast(aa, h=12), axes=F, main="ARIMA drift forecasts")
+lines(swpc)
+
+summary(aa)
+summary(forecast(aa, h=12))
+
+f <- forecast(aa, h=12)
+fore <- as.data.frame(f)
+
+#starting dates:
+sdates <- as.Date(paste0(s$Month,"-01"))
+newdates <- seq(sdates[length(sdates)]+31, length.out=12, by="month")
+
+sdf <- s
+sdf$type <- "measured"
+sdf$Month <- sdates
+
+new <- data.frame(Month=newdates, median=fore$`Point Forecast`, 
+           type="forecast")
+pdf <- rbind(sdf, new)
+
+vars <- data.frame(Month=new$Month, 
+                   l80=fore$`Lo 80`,
+                   l95=fore$`Lo 95`,
+                   h80=fore$`Hi 80`,
+                   h95=fore$`Hi 95`)
+                   
+afit <- data.frame(Month=sdf$Month, median=as.vector(fitted(aa)))
+
+pdf$type <- factor(pdf$type, levels=c("measured", "forecast"))
+
+## hacky ggplot abuse
+pdf("plots/forecast.pdf", 5, 3)
+ggplot(pdf, aes(x=Month, y=median/1e3)) +
[…]

### 2014-09-30 R/clusterPrices.R

@@ -31,11 +31,57 @@ image(log10(as.matrix(hm[,-1])+1))
 library("gplots")
 library("RColorBrewer")
 
+num.cols=24
+
 row.lab <- rep("", ncol(hm)-1)
 row.lab[seq(1, ncol(hm)-1, by=4)] <- colnames(hm)[-1][seq(1, ncol(hm)-1, by=4)]
-bl <- colorRampPalette(brewer.pal(9, "Blues"))(64)
+bl <- colorRampPalette(brewer.pal(9, "Blues"))(num.cols)
 # uneven breaks, norm distributed prices
+rsc <- as.numeric(factor(gsub("(.*?)\\d+", "\\1", hm[,1])))
+rsc <- rsc[complete.cases(hm)]
+
 heatmap.2(log10(as.matrix(hm[complete.cases(hm),-1])), 
           dendrogram="row", trace="none", Colv=NULL,
-          col=bl, labRow=F, labCol=row.lab, srtCol=45)
+          col=bl, labRow=F, labCol=row.lab, srtCol=45,
+          RowSideColors=heat.colors(max(rsc))[rsc],
+          breaks=quantile(log10(houses$Price), seq(0.01, .97, length.out=(num.cols+1))))
 
+length(rsc)
+
+library("dtw")
+ts.dist <- function(x, ...)
+  dist(x, method="DTW")
+
+
+# pre-compute some stuffs
+mat <- log10(as.matrix(hm[complete.cases(hm),-1]))
+rowcols <- heat.colors(max(rsc))[rsc]
+brks <- quantile(log10(houses$Price), seq(0.01, .97, length.out=(num.cols+1)))
+
+heatmap.2(mat, dendrogram="row", trace="none", Colv=NULL,
+          distfun=ts.dist, hclustfun=hc, RowSideColors=rowcols,
+          col=bl, labRow=F, labCol=row.lab, srtCol=45, breaks=brks)
+
+library("TSclust")
+sax.dist <- function(x, ...)
+  # try dist on un-transformed values
+  diss(as.matrix(hm[complete.cases(hm),-1]), "DWT", ...)
+
+hc <- function(x, ...)
+  hclust(x, method="ward.D")
+
+heatmap.2(mat, dendrogram="row", trace="none", Colv=NULL,
+          distfun=sax.dist, hclustfun=hc, RowSideColors=rowcols,
+          col=bl, labRow=F, labCol=row.lab, srtCol=45, breaks=brks)
+
[…]

### 2014-09-30 R/functions.R

@@ -85,9 +85,13 @@ theme_sdc_null <- function(base_size=12, base_family="sans") {
     panel.grid.major = element_line(size=.4),
     panel.grid.major.x = element_line(),
     panel.grid.major.y = element_line(),
-    panel.grid.minor = element_line(),
-    panel.grid.minor.x = element_line(size=.1),
-    panel.grid.minor.y = element_line(size=.1),
+    # removed:
+    #panel.grid.minor = element_line(),
+    #panel.grid.minor.x = element_line(size=.1),
+    #panel.grid.minor.y = element_line(size=.1),
+    panel.grid.minor = element_blank(),
+    panel.grid.minor.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
     ## transparent bg?
     plot.background = element_rect(fill = "transparent", colour = NA),
     plot.title = element_text(color="grey60", face=2, size=20),

### 2014-09-30 arima2.R

@@ -1,0 +1,164 @@
+## ARIMA forecasting
+library("dplyr")
+library("forecast")
+library("ggplot2")
+library("gridExtra")
+library("zoo")
+source("R/functions.R")
+
+houses <- loadHouses()
+
+## Background work, time series munging and tests
+s <- sw184
+s <- group_by(s, Month) %>% summarise(median=median(Price))
+
+# n.b. non-stationary time series
+
+# check for (partial) auto-correlation, look at detrended etc.
+plot(swpc, lag(swpc, 12))
+
+# no detectable periodicity in monthly data
+decompose(swpc)
+seasadj(stl(swpc))
+
+# residuals OK; pacf, acf PASS
+tsdisplay(diff(swpc))
+
+# positive reasonably linear trend, can be detrended
+m <- lm(coredata(swpc) ~ index(swpc))
+detr <- zoo(resid(m), index(swpc))
+plot(swpc)
+plot(detr)
+decompose(detr)
+
+## ARIMA model fitting and forecasting
+swpc <- ts(s$median, as.numeric(as.yearmon(s$Month)))
+
+aa <- auto.arima(swpc)
+plot(swpc)
+lines(fitted(aa), col="blue")
+plot(forecast(aa, h=12), axes=F, main="ARIMA drift forecasts")
+lines(swpc)
+
+summary(aa)
+summary(forecast(aa, h=12))
+
+f <- forecast(aa, h=12)
+fore <- as.data.frame(f)
+
+#starting dates:
+sdates <- as.Date(paste0(s$Month,"-01"))
+newdates <- seq(sdates[length(sdates)]+31, length.out=12, by="month")
+
+sdf <- s
+sdf$type <- "measured"
+sdf$Month <- sdates
+
+new <- data.frame(Month=newdates, median=fore$`Point Forecast`, 
+                  type="forecast")
+pdf <- rbind(sdf, new)
+
+vars <- data.frame(Month=new$Month, 
+                   l80=fore$`Lo 80`,
+                   l95=fore$`Lo 95`,
+                   h80=fore$`Hi 80`,
+                   h95=fore$`Hi 95`)
+
+afit <- data.frame(Month=sdf$Month, median=as.vector(fitted(aa)))
+
+pdf$type <- factor(pdf$type, levels=c("measured", "forecast"))
+
+## hacky ggplot abuse
+pdf("plots/forecast.pdf", 5, 3)
+ggplot(pdf, aes(x=Month, y=median/1e3)) +
[…]

### 2014-10-01 report/article_4.tex
begin latex report

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% LaTeX Template
% Version 1.1 (10/6/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
10pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
%twoside, % Two page layout (page indentation for binding and different headers)
headinclude,footinclude, % Extra spacing for the header and footer
BCOR5mm, % Binding correction
]{scrartcl}

\input{structure.tex} % Include the structure.tex file which specified the document structure and layout

\hyphenation{Fortran hy-phen-ation} % Specify custom hyphenation points in words with dashes where you would like hyphenation to occur, or alternatively, don't put any dashes in a word to stop hyphenation altogether

%% additional packages:
\usepackage{wrapfig}
\usepackage[left=1.5in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{caption}

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------

\title{\includegraphics[width=5cm]{Figures/datarea.pdf} \\
{\large \#SummerDataChallenge}} % The article title

\author{ {\large Benjamin L. Moore\textsuperscript{1}}} % The article author(s) - author affiliations need to be specified in the AUTHOR AFFILIATIONS block

\date{} % An optional date to appear under the author(s)

%----------------------------------------------------------------------------------------

% custom macros:

\newcommand*{\logo}{\includegraphics[scale=.22]{Figures/datarea.pdf}}

\begin{document}

%----------------------------------------------------------------------------------------
%	HEADERS
%----------------------------------------------------------------------------------------

\renewcommand{\sectionmark}[1]{\markright{\spacedlowsmallcaps{summerdatachallenge}}} % The header for all pages (oneside) or for even pages (twoside)
%\renewcommand{\subsectionmark}[1]{\markright{\thesubsection~#1}} % Uncomment when using the twoside option - this modifies the header on odd pages
\lehead{\mbox{\llap{\small\thepage\kern1em\color{halfgray} \vline}\color{halfgray}\hspace{0.5em}\rightmark\hfil}} % The header style

\pagestyle{scrheadings} % Enable the headers specified in this block

%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS & LISTS OF FIGURES AND TABLES
%----------------------------------------------------------------------------------------

\maketitle % Print the title/author/date block

\setcounter{tocdepth}{2} % Set the depth of the table of contents to show sections and subsections only

% \tableofcontents % Print the table of contents

% \listoffigures % Print the list of figures

% \listoftables % Print the list of tables

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------
\vspace{-4em}
\section*{Overview}

I analysed the {\bf London house prices} dataset is several innovative
ways. Firstly \emph{fractal context} shows how house sales within an
arbitrary area relate to those if neighbouring districs, up and down
the postcode hierarchy. Secondly, I employed \emph{ARIMA forecasting}
to extrapolate an area's price trend with an eye to suggesting
profitable investments. Finally, these and other metrics were combined
into an \emph{investment grade} for a given postcode area. Combined
this set of analyses forms the basis of \logo, a product aimed at
``democratising real estate investment'', offering individuals and
businesses quantitative insights for data-driven investment
decisions.

A fully-featured report is online at {\bf \href{http://blm.io/datarea}{blm.io} }and
scripts to reproduce all analyses are available from
{\bf \href{http://github.com/blmoore/summerdatachallenge}{github}}.

%----------------------------------------------------------------------------------------
%	AUTHOR AFFILIATIONS
%----------------------------------------------------------------------------------------

{\let\thefootnote\relax\footnotetext{\textsuperscript{1} \textit{MRC
      Human Genetics Unit, University of Edinburgh, Scotland, United Kingdom}}}

%----------------------------------------------------------------------------------------

%\newpage % Start the article content on the second page, remove this if you have a longer abstract that goes onto the second page

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\section{Description of analyses}

When considering a property in a given area, you may ask a realtor
questions like, how do prices here compare to surrounding postcodes?
How have prices increased over time?

\setlength{\intextsep}{.1em}
\begin{wrapfigure}{R}{.45\textwidth}
\begin{center}
\includegraphics[width=.44\textwidth]{Figures/fractal.png}
\caption{ Fractal context of postcode SW18 4HU in Wandsworth, South London.}
\end{center}
\end{wrapfigure}

A real estate professional can give an opinion based on anecdotal
evidence and the limited sample size they have experienced, but with a
large dataset we can address these questions quantitatively and
present them as clear visualisations.

\subsection{Fractal context}

Letting and sales sites may currently list some recent sales, but it's
not currently possible to see a quantitative overview of property
prices in a region, within the context of its sector, district and
postcode.

These questions can be answered with a clear and intuitive data
visualisation of price distributions in the postcode hierarchy --- we
call this the fractal context of a property price.

\subsection{ARIMA forecasting}

Given area price histories, it's of interest to investors to speculate
on potential investment returns for a given property area. This can be
done through time series analysis using the autoregressive integrated
moving average (ARIMA) model. Investment returns can be maximised by
selecting areas with maximal growth
projections.\textsuperscript{$\star$} \\

\begin{figure}[h]
\begin{center}
\includegraphics[width=.9\textwidth]{Figures/arima.png}
\caption{Example ARIMA forecast of median house prices for a specific postcode.}
\end{center}
\end{figure}

\vspace{-2em}
\subsection{Investment grading}

Combining growth forecasts and contextual information, as well as
other derived metrics from this dataset, allow an approximate
investment grading, relative to other postcodes within the area
covered by this dataset. This intuitive output metric can help
prioritise areas ripe for property investment. \\

{\let\thefootnote\relax\footnotetext{\textsuperscript{$\star$}
    Data intended to assist investors and does not constitute
    investment advice; independent advice should be sought where appropriate. }}

%----------------------------------------------------------------------------------------
%	METHODS
%----------------------------------------------------------------------------------------
\vspace{-2em}
\section{Value generation}

\logo \hspace{.1em} would first be released as both a web service and
associated apps for mobiles and tablets. These would offer limited
analysis and visualisation for free, with user analytics and feedback
helping to validate the business model.

\begin{wrapfigure}{r}{.4\textwidth}
\begin{center}
\includegraphics[width=.39\textwidth]{Figures/mockup.png}
\caption{ iOS app mock-up.}
\end{center}
\end{wrapfigure}

A small data science team could expand analyses, integrating novel
public and privately-acquired datasets and release them under a
rolling subscription model, aimed at private landlords and property
speculators, as well as professional real estate investment trusts
(REITs). Significant revenue could be generated through tailored
partnerships with funds and high net-worth individuals.

Should the product not succeed in a direct-to-consumer market, the
business could pivot to an analytics provider for existing letting and
sales sites, offering unique data insights indirectly to the consumer
market, or as independent fund consultants for the financial
services. Having produced a functional product positions \logo \hspace{.1em} to
succeed in innovative real estate analytics.

\begin{flushright}
{\small Full version online at: \href{http://blm.io/datarea}{blm.io/datarea}}
\end{flushright}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=.1\textwidth]{Figures/logo.png} \\
\includegraphics[scale=.4]{Figures/datarea.pdf} \\
\caption*{ {\sf Democratising real estate investment}}
\end{center}
\end{figure}

\end{document}

### 2014-10-01 report/article_4.tex
logo swap

@@
 \usepackage{wrapfig}
-\usepackage[left=1.5in,right=1in,top=1in,bottom=1in]{geometry}
+\usepackage[left=1.2in,right=1in,top=1in,bottom=1in]{geometry}
 \usepackage{caption}
@@
 
-\title{\includegraphics[width=5cm]{Figures/datarea.pdf} \\
+\title{ \vspace{-3em}
+\includegraphics[width=5cm]{Figures/datarea.pdf} \\
 {\large \#SummerDataChallenge}} % The article title
@@
 ``democratising real estate investment'', offering individuals and
-businesses quantitative insights for data-driven investment
+businesses quantitative real estate insights for data-driven investment
 decisions.
 
-A fully-featured report is online at {\bf \href{http://blm.io/datarea}{blm.io} }and
-scripts to reproduce all analyses are available from
-{\bf \href{http://github.com/blmoore/summerdatachallenge}{github}}.
+A full report with additional visualisations is online at {\bf
+  \href{http://blm.io/datarea}{blm.io/datarea} }and scripts to
+reproduce all analyses are available from {\bf
+  \href{http://github.com/blmoore/summerdatachallenge}{github.com/blmoore/summerdatachallenge}}.
 
@@
 When considering a property in a given area, you may ask a realtor
-questions like, how do prices here compare to surrounding postcodes?
-How have prices increased over time?
+questions like: how do prices here compare to surrounding postcodes?
+How have prices increased over time?  A real estate professional can
+give an opinion based on their own limited sample size, but with a
+large dataset we can address these questions quantitatively and
+visualise the results.
 
-\setlength{\intextsep}{.1em}
+\subsection{Fractal context}
+
+\setlength{\intextsep}{0em}
 \begin{wrapfigure}{R}{.45\textwidth}
@@
 \end{wrapfigure}
-
-A real estate professional can give an opinion based on anecdotal
-evidence and the limited sample size they have experienced, but with a
-large dataset we can address these questions quantitatively and
-present them as clear visualisations.
-
-\subsection{Fractal context}
 
@@
 visualisation of price distributions in the postcode hierarchy --- we
-call this the fractal context of a property price.
+call this the fractal context of a property price (Figure 1).
 
-\subsection{ARIMA forecasting}
+\subsection{ARIMA modelling}
 
-Given area price histories, it's of interest to investors to speculate
-on potential investment returns for a given property area. This can be
-done through time series analysis using the autoregressive integrated
-moving average (ARIMA) model. Investment returns can be maximised by
-selecting areas with maximal growth
-projections.\textsuperscript{$\star$} \\
+Given a time series of housing prices, it's of interest to investors
+to speculate on potential investment returns for a given property
+area. This can be done through time series analysis using the
+autoregressive integrated moving average (ARIMA) model (Figure
+2). Investment returns can be optimised by selecting areas with
+maximal growth projections.\textsuperscript{$\star$} \\
 
@@
 
-A small data science team could expand analyses, integrating novel
-public and privately-acquired datasets and release them under a
-rolling subscription model, aimed at private landlords and property
-speculators, as well as professional real estate investment trusts
-(REITs). Significant revenue could be generated through tailored
-partnerships with funds and high net-worth individuals.
+A small data science team will gradually expand the range of analyses
+offered, integrating novel public and privately-acquired datasets and
+release them under a rolling subscription model, aimed at private
+landlords and property speculators, as well as professional real
+estate investment trusts (REITs). Significant revenue could be
+generated through tailored partnerships with funds and high net-worth
+individuals.
 
-Should the product not succeed in a direct-to-consumer market, the
-business could pivot to an analytics provider for existing letting and
-sales sites, offering unique data insights indirectly to the consumer
-market, or as independent fund consultants for the financial
-services. Having produced a functional product positions \logo \hspace{.1em} to
-succeed in innovative real estate analytics.
-
-\begin{flushright}
-{\small Full version online at: \href{http://blm.io/datarea}{blm.io/datarea}}
-\end{flushright}
+Additionally the business could pivot to an analytics provider for
+existing letting and sales sites, offering unique data insights
+indirectly to the consumer market, or as independent fund consultants
+for the financial services. Having produced a first-to-market
+innovative product positions \logo \hspace{.1em} to succeed as a
+provider of advanced real estate investment analytics. \\
 
@@
 
+\vspace{1em}
+\begin{flushright}
+{\small An enhanced overview of this entry is available online at: \href{http://blm.io/datarea}{blm.io/datarea}}
+\end{flushright}
+
 \end{document}

### 2014-10-02 report/article_4.tex
logo swap, colour change

@@
 \usepackage{wrapfig}
-\usepackage[left=1.2in,right=1in,top=1in,bottom=1in]{geometry}
+\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
 \usepackage{caption}
@@
 \title{ \vspace{-3em}
-\includegraphics[width=5cm]{Figures/datarea.pdf} \\
-{\large \#SummerDataChallenge}} % The article title
+\includegraphics[width=.1\textwidth]{Figures/logo.png} \\
+\includegraphics[width=4cm]{Figures/datarea.pdf} \\
+} % The article title
 
-\author{ {\large Benjamin L. Moore\textsuperscript{1}}} % The article author(s) - author affiliations need to be specified in the AUTHOR AFFILIATIONS block
+\author{  Summer data challenge entry by Benjamin
+  L. Moore\textsuperscript{1} } % The article author(s) - author affiliations need to be specified in the AUTHOR AFFILIATIONS block
 
@@
 arbitrary area relate to those if neighbouring districs, up and down
-the postcode hierarchy. Secondly, I employed \emph{ARIMA forecasting}
-to extrapolate an area's price trend with an eye to suggesting
-profitable investments. Finally, these and other metrics were combined
-into an \emph{investment grade} for a given postcode area. Combined
-this set of analyses forms the basis of \logo, a product aimed at
-``democratising real estate investment'', offering individuals and
-businesses quantitative real estate insights for data-driven investment
-decisions.
+the postcode hierarchy. Secondly, I fit \emph{ARIMA models} to
+forecast an area's price trend into the future, with an eye to
+suggesting profitable investments. Finally, these and other metrics
+were combined into an \emph{investment grade} for a given postcode
+area. Combined this set of analyses forms the basis of \logo, a
+product aimed at ``democratising real estate investment'', offering
+individuals and businesses quantitative real estate insights for
+data-driven investment decisions.
 
@@
 
-\section{Description of analyses}
+\section*{Description of analyses}
 
@@
 \begin{center}
-\includegraphics[width=.44\textwidth]{Figures/fractal.png}
+\includegraphics[width=.44\textwidth]{Figures/fractal.pdf}
 \caption{ Fractal context of postcode SW18 4HU in Wandsworth, South London.}
@@
 \begin{center}
-\includegraphics[width=.9\textwidth]{Figures/arima.png}
+\includegraphics[width=.9\textwidth]{Figures/arima.pdf}
 \caption{Example ARIMA forecast of median house prices for a specific postcode.}

### 2014-10-02 R/arima.R

@@ -3,8 +3,10 @@ library("dplyr")
 library("forecast")
 library("ggplot2")
 library("zoo")
+source("R/functions.R")
 
-s <- sw184[sw184$Postcode == "SW18 4HU",]
+houses <- loadHouses()
+s <- houses[houses$Postcode == "SW18 4HU",]
 s <- group_by(s, Month) %>% summarise(median=median(Price))
 
 # n.b. non-stationary time series
@@ -77,7 +79,7 @@ ggplot(pdf, aes(x=Month, y=median/1e3)) +
               alpha=I(.2)) +
   geom_ribbon(data=vars, inherit.aes=F,
               aes(x=Month, ymin=l95/1e3, ymax=h95/1e3),
-              alpha=I(.1)) +
+              alpha=I(.1), fill=I(rgb(61, 133, 193, max=255))) +
   labs(y="Median house price (£k)", x="") +
   theme(legend.position=c(.2,.8)) +
   annotate("text", x=as.Date("2010-06-01"), size=8,

### 2014-10-02 arima2.R

@@ -9,10 +9,11 @@ source("R/functions.R")
 houses <- loadHouses()
 
 ## Background work, time series munging and tests
-s <- sw184
+s <- houses[houses$sector == "SW18 4",]
 s <- group_by(s, Month) %>% summarise(median=median(Price))
 
 # n.b. non-stationary time series
+swpc <- ts(s$median, as.numeric(as.yearmon(s$Month)))
 
 # check for (partial) auto-correlation, look at detrended etc.
 plot(swpc, lag(swpc, 12))
@@ -69,20 +70,22 @@ afit <- data.frame(Month=sdf$Month, median=as.vector(fitted(aa)))
 pdf$type <- factor(pdf$type, levels=c("measured", "forecast"))
 
 ## hacky ggplot abuse
-pdf("plots/forecast.pdf", 5, 3)
+pdf("plots/forecast_v2.pdf", 5, 3)
 ggplot(pdf, aes(x=Month, y=median/1e3)) +
-  geom_point(data=subset(pdf, type=="measured")) +
-  geom_line(data=subset(pdf, type=="forecast")) + theme_sdc() +
   #scale_color_manual() +
-  # arima model fit
-  geom_line(data=afit, col=I("grey40"), linetype=2) +
   # forecase bounds
   geom_ribbon(data=vars, inherit.aes=F,
-              aes(x=Month, ymin=l80/1e3, ymax=h80/1e3),
-              alpha=I(.2)) +
-  geom_ribbon(data=vars, inherit.aes=F,
               aes(x=Month, ymin=l95/1e3, ymax=h95/1e3),
-              alpha=I(.1)) +
+              alpha=I(.5), fill=I(rgb(192, 213, 237, max=255))) +
+  geom_ribbon(data=vars, inherit.aes=F,
+              aes(x=Month, ymin=l80/1e3, ymax=h80/1e3),
+              alpha=I(.5), fill=I(rgb(61, 133, 193, max=255))) +
+  geom_line(data=subset(pdf, type=="forecast")) + theme_sdc() +
+  # arima model fit
+  geom_line(data=afit, col=I(rgb(192, 213, 237, max=255)), 
+            linetype=1, size=3.5, alpha=I(.5)) +
+  geom_line(data=afit, col=I(rgb(61, 133, 193, max=255)), linetype=2) +
+  geom_point(data=subset(pdf, type=="measured"), col=I("grey40")) +
   labs(y="Median house price (£k)", x="") +
[…]

### 2014-10-04 report/article_4.tex
investment grading

@@
 \includegraphics[width=.1\textwidth]{Figures/logo.png} \\
-\includegraphics[width=4cm]{Figures/datarea.pdf} \\
+\includegraphics[width=4cm]{Figures/logotext.png} \\
 } % The article title
@@
 
-\newcommand*{\logo}{\includegraphics[scale=.22]{Figures/datarea.pdf}}
+\newcommand*{\logo}{\includegraphics[scale=.04]{Figures/logotext.png}}
 
@@
 
-\section*{Description of analyses}
+\vspace{-1em}
+\section*{Analyses}
 
@@
 
-\subsection{Fractal context}
+\subsection*{Fractal context}
 
@@
 \begin{center}
-\includegraphics[width=.44\textwidth]{Figures/fractal.pdf}
+\includegraphics[width=.44\textwidth]{Figures/fractal.png}
 \caption{ Fractal context of postcode SW18 4HU in Wandsworth, South London.}
@@
 
-\subsection{ARIMA modelling}
+A specific postcode or area can be marked within its distribution,
+giving a clear view of a properties relative pricing which cannot be
+gleaned just by browsing recent sales. The example view (Figure 1)
+combines ranked violin density plots and a geo-heatmap for a clear,
+intuitive view of a locations price context.
+
+\subsection*{ARIMA modelling}
 
@@
 \begin{center}
-\includegraphics[width=.9\textwidth]{Figures/arima.pdf}
+\includegraphics[width=.9\textwidth]{Figures/arima.png}
 \caption{Example ARIMA forecast of median house prices for a specific postcode.}
@@
 \vspace{-2em}
-\subsection{Investment grading}
+\subsection*{Investment grading}
 
-Combining growth forecasts and contextual information, as well as
-other derived metrics from this dataset, allow an approximate
-investment grading, relative to other postcodes within the area
-covered by this dataset. This intuitive output metric can help
-prioritise areas ripe for property investment. \\
+Combining growth forecasts and other derived metrics from this
+dataset allow an approximate investment grading, relative to other
+postcodes within the area covered by this dataset. This intuitive
+output metric can help prioritise areas ripe for property investment.
+
+As an example, Figure 3 shows the top 5 best-ranked postcode sectors
+in which to invest, according to their 12-month price growth forecast
+and historical annualised volatility --- thus combining the empirical
+data with theoretical model outputs. \\
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=.9\textwidth]{Figures/t5.pdf}
+\caption{Top 5 property investment sectors by projected returns and historical volatility.}
+\end{center}
+\end{figure}
+
+The rankings used in Figure 3 made use of the $drift$ parameter of
+ARIMA models, combined with statistical volatility over the five year
+house prices dataset. Both parameters were converted to quantiles and
+weighted equally, with those sectors with the largest projected growth
+and lowest historical volatility receive the highest investment grade.
+
+\vspace{-1em}
+\section*{Summary}
+These initial
 
@@
 %----------------------------------------------------------------------------------------
-\vspace{-2em}
-\section{Value generation}
+\clearpage
+\section*{Value generation}
 
@@
 \includegraphics[width=.1\textwidth]{Figures/logo.png} \\
-\includegraphics[scale=.4]{Figures/datarea.pdf} \\
+\includegraphics[scale=.12]{Figures/logotext.png} \\
 \caption*{ {\sf Democratising real estate investment}}

### 2014-10-04 R/grading.R

@@ -1,0 +1,114 @@
+# ##########################################################
+# ----------------- Investment Grading ------------------- #
+#                                                          #
+# Pre-compute stats for all sectors:                       #
+#                                                          #
+#   1) growth:     signed % change from start of forecast  #
+#                  to end (12 months later)                #
+#   2) volatility: how valitile has the price been over    #
+#                  the last 5 years                        #
+#                                                          #
+# Find all as a quantile of all sectors, mean could then   #
+# be its "investment grade" depending on its distribution. #
+############################################################
+library("dplyr")
+library("forecast")
+library("ggplot2")
+library("parallel")
+library("zoo")
+source("R/functions.R")
+
+houses <- loadHouses()
+
+growth <- function(sector){
+  data <- houses[houses$sector == sector,]
+  data <- group_by(data, Month) %>% summarise(median=median(Price))
+  data.ts <- ts(data$median, as.numeric(as.yearmon(data$Month)))
+  arima <- auto.arima(data.ts, allowdrift=T)
+#  f <- as.data.frame(forecast(arima, h=12))
+  #s.e <- fore$`Point Forecast`[c(1, 12)]
+  
+  # percentage change over forecast
+  #100*((s.e[2] - s.e[1]) / s.e[1])
+ # message(coef(arima))
+  if("drift" %in% names(coef(arima))){
+    return(coef(arima)[["drift"]])
+  } else {
+    return(0L)
+  }
+}
+
+growth <- Vectorize(growth)
+
+# cut out some secots with too few datapoints
+secs <- group_by(houses, sector) %>% 
+  summarise(counts=n()) %>% filter(counts > 500)
+
+df <- data.frame(sector=secs$sector)
+
+## takes ~ 5 mins to run on 8 cores
+df$growth <- pvec(as.vector(df$sector), growth, mc.cores=8)
+
+## convert raw growth vals to quantiles
+df$growth.q <- ecdf(df$growth)(df$growth)
+
+hist(df$growth.q)
+
+tc <- houses[houses$sector == "NW9 6",]
+plot(tc$Trdate, tc$Price)
[…]

### 2014-10-04 arima2.R

@@ -124,6 +124,7 @@ plotForecast <- function(out){
 modelHPs <- function(sector){
   #sector="W6 7"
   data <- houses[houses$sector == sector,]
+  # aggregate by monht
   data <- group_by(data, Month) %>% summarise(median=median(Price))
   data.ts <- ts(data$median, as.numeric(as.yearmon(data$Month)))
   arima <- auto.arima(data.ts)

### 2014-10-06 report/article_4.tex
report update

@@
 questions like: how do prices here compare to surrounding postcodes?
-How have prices increased over time?  A real estate professional can
-give an opinion based on their own limited sample size, but with a
-large dataset we can address these questions quantitatively and
-visualise the results.
+How have prices increased over time? And how have prices changed over
+time? A real estate professional can give an opinion based on their
+own limited sample size, but with a large dataset we can address these
+questions quantitatively and visualise the results.
 
@@
 \setlength{\intextsep}{0em}
-\begin{wrapfigure}{R}{.45\textwidth}
+\begin{wrapfigure}{R}{.5\textwidth}
+\vspace{1em}
 \begin{center}
-\includegraphics[width=.44\textwidth]{Figures/fractal.png}
+\includegraphics[width=.48\textwidth]{Figures/fractal.png}
+\vspace{.5em}
 \caption{ Fractal context of postcode SW18 4HU in Wandsworth, South London.}
@@
 area. This can be done through time series analysis using the
-autoregressive integrated moving average (ARIMA) model (Figure
-2). Investment returns can be optimised by selecting areas with
-maximal growth projections.\textsuperscript{$\star$} \\
+autoregressive integrated moving average (ARIMA) model (Figure 2). The
+model fitting procedure included regularisation that allowed a $drift$
+term to model non-stationary trends, as well as coefficients capturing
+periodicity or seasonal effects. Potential investment returns could be
+optimised by selecting areas with
+maximal growth projections. \\
 
@@
 
-\vspace{-2em}
+\vspace{-1em}
 \subsection*{Investment grading}
@@
 and historical annualised volatility --- thus combining the empirical
-data with theoretical model outputs. \\
+data with theoretical model outputs. These include South London
+suburbs of Upper Tooting and Morden, Walthamstow, as well as Brockley,
+whose train station within SE4 1 was linked with the London overground network in 2010.\\
 
@@
 weighted equally, with those sectors with the largest projected growth
-and lowest historical volatility receive the highest investment grade.
-
-\vspace{-1em}
-\section*{Summary}
-These initial
+and lowest historical volatility receive the highest investment grade.\textsuperscript{$\star$}
 
@@
 
+\subsection*{Socio-economic value}
+
+\subsection*{Initial product}
 \logo \hspace{.1em} would first be released as both a web service and
@@
 
+\subsection*{Business development}
+
 \vspace{1em}

### 2014-10-06 report/article_4.tex
writeup devel

@@
 area. Combined this set of analyses forms the basis of \logo, a
-product aimed at ``democratising real estate investment'', offering
+startup aimed at ``democratising real estate investment'', offering
 individuals and businesses quantitative real estate insights for
@@
 %----------------------------------------------------------------------------------------
-\clearpage
+%\clearpage
 \section*{Value generation}
 
-\subsection*{Socio-economic value}
+The above-described analyses form the basis of \logo\hspace{.1em}, a
+start-up aimed at ``democratising real estate investment'' by
+making deep-dive analytics data available both to a consumer and
+professional investment market.
 
-\subsection*{Initial product}
-\logo \hspace{.1em} would first be released as both a web service and
-associated apps for mobiles and tablets. These would offer limited
-analysis and visualisation for free, with user analytics and feedback
-helping to validate the business model.
+\subsection*{1. Analytics provider for online house sales and lettings
+agents}
+
+\logo\hspace{.1em} can provide {\bf real-time property area analytics} to
+existing market leaders in {\bf online sales and lettings}.
+
+\begin{wrapfigure}{l}{.2\textwidth}
+\begin{center}
+\includegraphics[width=.15\textwidth]{Figures/zoopla.png}
+\includegraphics[width=.2\textwidth]{Figures/rightmove.jpg}
+\caption*{ Example customer segments.}
+\end{center}
+\end{wrapfigure}
+
+There are currently no mainsteam property sales or rental sites that
+attempts to provide any quantitative insights to contextualise a
+property's list price. Some list recent nearby sales in plain text,
+but the wealth of historical sales data available allows for much
+richer displays (as demonstrated with the fractal context view,
+\emph{above}), which offer valuable and practical insights to
+prospective homebuyers.
+
+This added feature would help differentiate a given sales and letting
+site as well as increase the brand positioning of \logo\hspace{.1em}
+within the real estate sector. Contextual maps could be drawn as
+interactive D3.js visualisations and neatly integrated with a typical
+property search. Additionally, web analytics data from our partnered
+site could be fed-back to help further enhancement and expansion of
+the data-driven insights and visualisations we offer.
+
+\subsection*{2. Direct-to-consumer demo application}
+\logo \hspace{.1em} would release both a {\bf web service} and associated
+{\bf mobiles and tablet applications}. These would offer limited analysis and
+visualisation without any fee, with collected user data and feedback helping to
+validate the business model and develop our software.
 
@@
 \caption{ iOS app mock-up.}
+\vspace{-2em}
 \end{center}
@@
 
-Additionally the business could pivot to an analytics provider for
-existing letting and sales sites, offering unique data insights
-indirectly to the consumer market, or as independent fund consultants
-for the financial services. Having produced a first-to-market
-innovative product positions \logo \hspace{.1em} to succeed as a
-provider of advanced real estate investment analytics. \\
+\subsection*{3. Real estate data science consultancy}
 
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=.1\textwidth]{Figures/logo.png} \\
-\includegraphics[scale=.12]{Figures/logotext.png} \\
-\caption*{ {\sf Democratising real estate investment}}
-\end{center}
-\end{figure}
+Having established a property area analytics brand, \logo\hspace{.1em}
+will look to develop business relationships with real estate
+investment funds and high net-worth individuals as a data-driven
+propert investment consultancy and independent analyst for the
+financial services. Our domain expertise and existing partnerships and
+products would place us as a market leader in this undeveloped segment.
 
-\subsection*{Business development}
+% \begin{figure}[h!]
+% \begin{center}
+% \includegraphics[width=.1\textwidth]{Figures/logo.png} \\
+% \includegraphics[scale=.12]{Figures/logotext.png} \\
+% \caption*{ {\sf Democratising real estate investment}}
+% \end{center}
+% \end{figure}
 

### 2014-10-08 report/article_4.tex
report final for today
save report before latex env swap

@@
 
+\setlength{\intextsep}{0em}
+\begin{wrapfigure}{R}{.3\textwidth}
+%\vspace{-1em}
+\centering
+\includegraphics[width=.28\textwidth]{Figures/overview.png}
+%\vspace{-1em}
+\caption{ Mapped house sales, brightness is proportional to price.}
+\end{wrapfigure}
+
 I analysed the {\bf London house prices} dataset is several innovative
@@
 A full report with additional visualisations is online at {\bf
-  \href{http://blm.io/datarea}{blm.io/datarea} }and scripts to
+  \leavevmode\href{http://blm.io/datarea}{blm.io/datarea} }and scripts to
 reproduce all analyses are available from {\bf
-  \href{http://github.com/blmoore/summerdatachallenge}{github.com/blmoore/summerdatachallenge}}.
+  \leavevmode\href{http://github.com/blmoore/summerdatachallenge}{github.com/blmoore/summerdatachallenge}}.
 
@@
 
+%\setlength{\intextsep}{0em}
+\begin{wrapfigure}[17]{L}{.45\textwidth}
+%\vspace{-1em}
+\centering
+\includegraphics[width=.43\textwidth]{Figures/fractal.png}
+\caption{ Fractal context of postcode SW18 4HU in Wandsworth, South London.}
+%\vspace{35em}
+\end{wrapfigure}
+
 \subsection*{Fractal context}
-
-\setlength{\intextsep}{0em}
-\begin{wrapfigure}{R}{.5\textwidth}
-\vspace{1em}
-\begin{center}
-\includegraphics[width=.48\textwidth]{Figures/fractal.png}
-\vspace{.5em}
-\caption{ Fractal context of postcode SW18 4HU in Wandsworth, South London.}
-\end{center}
-\end{wrapfigure}
 
@@
 
-\begin{wrapfigure}{l}{.2\textwidth}
+\begin{wrapfigure}{l}{.25\textwidth}
 \begin{center}
-\includegraphics[width=.15\textwidth]{Figures/zoopla.png}
-\includegraphics[width=.2\textwidth]{Figures/rightmove.jpg}
-\caption*{ Example customer segments.}
+\includegraphics[width=.18\textwidth]{Figures/zoopla.png}
+\includegraphics[width=.23\textwidth]{Figures/rightmove.jpg}
+\caption*{ Potential partners. }
 \end{center}
@@
 \begin{flushright}
-{\small An enhanced overview of this entry is available online at: \href{http://blm.io/datarea}{blm.io/datarea}}
+{\small An enhanced overview of this entry is available online at: \leavevmode\href{http://blm.io/datarea}{blm.io/datarea}}
 \end{flushright}

### 2014-10-08 R/report_viz.R

@@ -1,0 +1,23 @@
+source("R/functions.R")
+
+houses <- loadHouses()
+
+
+library("ggmap")
+map <- get_map(location="London", zoom=8, maptype="toner", color="bw", source="stamen")
+p1 <- ggmap(map, extent="device", darken=.6)
+p1 + geom_point(data=houses[sample(1:nrow(houses), 1e5),], 
+                aes(x=Longitude, y=Latitude, alpha=log10(Price)), col="white") +
+  scale_alpha_continuous(range=c(.01, .2)) +
+  labs(x="", y="") + theme(legend.position="none") 
+
+p1+ geom_point(data=houses[houses$yearmon == levels(houses$yearmon)[ym],], 
+                  aes(x=Longitude, y=Latitude, alpha=log10(log10(Price))), 
+                  col="white", size=.3) + 
+    labs(x="", y="") + 
+    theme(legend.position="none",
+          axis.ticks = element_blank(), axis.text.x = element_blank(),
+          axis.text.y = element_blank()) +
+    annotate("text", x=.8, y=50.5, label=levels(houses$yearmon)[ym], 
+             col="white", family="mono") +
+    scale_alpha_continuous(range=c(.2, .45))

### 2014-10-08 report/article_4.tex
manual placement

@@
 
+\subsection*{Fractal context}
+
+Letting and sales sites may currently list some recent sales, but it's
+not currently possible to see a
+
 %\setlength{\intextsep}{0em}
-\begin{wrapfigure}[17]{L}{.45\textwidth}
-%\vspace{-1em}
+\begin{wrapfigure}[16]{l}{.45\textwidth}
+\vspace{.5em}
 \centering
@@
 
-\subsection*{Fractal context}
-
-Letting and sales sites may currently list some recent sales, but it's
-not currently possible to see a quantitative overview of property
+\noindent quantitative overview of property
 prices in a region, within the context of its sector, district and
@@
 
-\begin{wrapfigure}{r}{.4\textwidth}
-\begin{center}
+\begin{wrapfigure}[11]{r}{.4\textwidth}
+\centering
 \includegraphics[width=.39\textwidth]{Figures/mockup.png}
@@
 \vspace{-2em}
-\end{center}
 \end{wrapfigure}

### 2014-10-09 report/article_4.tex
new report header

@@
 
-\renewcommand{\sectionmark}[1]{\markright{\spacedlowsmallcaps{summerdatachallenge}}} % The header for all pages (oneside) or for even pages (twoside)
+%\renewcommand{\sectionmark}[1]{
+\markright{\spacedlowsmallcaps{summerdatachallenge}}
+%} % The header for all pages (oneside) or for even pages (twoside)
 %\renewcommand{\subsectionmark}[1]{\markright{\thesubsection~#1}} % Uncomment when using the twoside option - this modifies the header on odd pages
@@
 
-\maketitle % Print the title/author/date block
+% \maketitle % Print the title/author/date block
 
@@
 % \listoftables % Print the list of tables
+
+% Make header in inkscape?
+\begin{figure}[width=\textwidth]
+\vspace{-3em}
+\centering
+\includegraphics[width=.7\textwidth]{Figures/header2.pdf}
+\vspace{.5em}
+\end{figure}
+
+%suppres headers
+\thispagestyle{plain}
 
@@
 \includegraphics[width=.28\textwidth]{Figures/overview.png}
-%\vspace{-1em}
-\caption{ Mapped house sales, brightness is proportional to price.}
+\vspace{-1em}
+\caption*{ A map of the house prices dataset (brightness is proportional to
+  price and density).}
 \end{wrapfigure}

### 2014-10-09 R/report_viz.R

@@ -1,23 +1,15 @@
+library("ggmap")
 source("R/functions.R")
+set.seed(42)
 
 houses <- loadHouses()
 
-
-library("ggmap")
 map <- get_map(location="London", zoom=8, maptype="toner", color="bw", source="stamen")
 p1 <- ggmap(map, extent="device", darken=.6)
-p1 + geom_point(data=houses[sample(1:nrow(houses), 1e5),], 
-                aes(x=Longitude, y=Latitude, alpha=log10(Price)), col="white") +
-  scale_alpha_continuous(range=c(.01, .2)) +
-  labs(x="", y="") + theme(legend.position="none") 
 
-p1+ geom_point(data=houses[houses$yearmon == levels(houses$yearmon)[ym],], 
-                  aes(x=Longitude, y=Latitude, alpha=log10(log10(Price))), 
-                  col="white", size=.3) + 
-    labs(x="", y="") + 
-    theme(legend.position="none",
-          axis.ticks = element_blank(), axis.text.x = element_blank(),
-          axis.text.y = element_blank()) +
-    annotate("text", x=.8, y=50.5, label=levels(houses$yearmon)[ym], 
-             col="white", family="mono") +
-    scale_alpha_continuous(range=c(.2, .45))
+pdf("plots/report_overview.pdf", 2.5, 2.5)
+p1 + geom_point(data=houses[sample(1:nrow(houses), 1e5),], size=I(.2),
+                aes(x=Longitude, y=Latitude, alpha=log10(Price)), col="white") +
+  scale_alpha_continuous(range=c(.05, .25)) +
+  labs(x="", y="") + theme(legend.position="none") 
+dev.off()

### 2014-10-09 report/article_4.tex
bringing web writeup inline with report

@@
 house prices dataset. Both parameters were converted to quantiles and
-weighted equally, with those sectors with the largest projected growth
-and lowest historical volatility receive the highest investment grade.\textsuperscript{$\star$}
+weighted equally; those sectors with the highest projected growth
+and least historical volatility receive the highest investment grade
+(on a scale of AAA, AA, A, BBB \ldots C).\textsuperscript{$\star$}
+
+Returning to our original example, we can conclude that SW18 4HU is a
+promising choice of area for property investment. It's approximately
+median-priced for the sector, which is the cheapest of SW18 (in turn
+part of the desirable SW London area; Figure 1). It has good growth
+projections of XX in the next 12 months (Figure 2), and when combined
+with volatility and compared with all $\approx2600$ sectors in the
+dataset, receives the top AAA investment grade, placing it in the
+90\textsuperscript{th} percentile for predicted price growth and low
+historical volatility.
 
@@
 %----------------------------------------------------------------------------------------
-%\clearpage
+\clearpage
 \section*{Value generation}
@@
 The above-described analyses form the basis of \logo\hspace{.1em}, a
-start-up aimed at ``democratising real estate investment'' by
-making deep-dive analytics data available both to a consumer and
-professional investment market.
+start-up aimed at ``democratising real estate investment'' by making
+deep-dive analytics data available both to a consumer and professional
+investment market. Initially \logo\hspace{.1em} is a provider of
+information aimed at investors but cannot dispense investment advice
+until authorised to do so by the FCA. Three initial areas of interest
+to commercialise and generate social and economic value are listed:
 
@@
 
-\begin{wrapfigure}[11]{r}{.4\textwidth}
+\begin{wrapfigure}[12]{r}{.4\textwidth}
 \centering
@@
 \caption{ iOS app mock-up.}
-\vspace{-2em}
+%\vspace{-1em}
 \end{wrapfigure}
+
+Releasing these products aligns with our aims of democratising real
+estate investment. Amateur property investors and prospective
+landlords can make data-driven decisions, the likes of which are likely
+already employed by large investment trusts and funds.
 
@@
 will look to develop business relationships with real estate
-investment funds and high net-worth individuals as a data-driven
-propert investment consultancy and independent analyst for the
+investment funds and high net-worth individuals as a {\bf quantitative
+  property investment consultancy} and independent analyst for the
 financial services. Our domain expertise and existing partnerships and
-products would place us as a market leader in this undeveloped segment.
+products would place us as a market leader in this undeveloped
+segment.
 
@@
 \vspace{1em}
-\begin{flushright}
-{\small An enhanced overview of this entry is available online at: \leavevmode\href{http://blm.io/datarea}{blm.io/datarea}}
+\begin{flushright} {\small An enhanced online version of this short
+    report is available at:
+    \leavevmode\href{http://blm.io/datarea}{blm.io/datarea}}
 \end{flushright}

### 2014-10-09 report/structure.tex
bringing web writeup inline with report

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% Structure Specification File
%
% This file has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	REQUIRED PACKAGES
%----------------------------------------------------------------------------------------

\usepackage[ nochapters, % Turn off chapters since this is an article
beramono, % Use the Bera Mono font for monospaced text (\texttt)
eulermath,% Use the Euler font for mathematics 
pdfspacing, % Makes use of pdftex’ letter spacing capabilities via the microtype package
dottedtoc % Dotted lines leading to the page numbers in the table
%          ofcontents 
]{classicthesis} % The layout is based on the Classic Thesis style

\usepackage{arsclassica} % Modifies the Classic Thesis package

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs

\usepackage[utf8]{inputenc} % Required for including letters with accents

\usepackage{graphicx} % Required for including images
\graphicspath{{Figures/}} % Set the default folder for images

\usepackage{enumitem} % Required for manipulating the whitespace between and within lists

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

\usepackage{subfig} % Required for creating figures with multiple parts (subfigures)

\usepackage{amsmath,amssymb,amsthm} % For including math equations, theorems, symbols, etc

\usepackage{varioref} % More descriptive referencing

%----------------------------------------------------------------------------------------
%	THEOREM STYLES
%---------------------------------------------------------------------------------------

\theoremstyle{definition} % Define theorem styles here based on the definition style (used for definitions and examples)
\newtheorem{definition}{Definition}

\theoremstyle{plain} % Define theorem styles here based on the plain style (used for theorems, lemmas, propositions)
\newtheorem{theorem}{Theorem}

\theoremstyle{remark} % Define theorem styles here based on the remark style (used for remarks and notes)

%----------------------------------------------------------------------------------------
%	HYPERLINKS
%---------------------------------------------------------------------------------------

\hypersetup{
%draft, % Uncomment to remove all links (useful for printing in black and white)
colorlinks=true, breaklinks=true, bookmarks=true,bookmarksnumbered,
urlcolor=webbrown, linkcolor=RoyalBlue, citecolor=webgreen, % Link colors
pdftitle={}, % PDF title
pdfauthor={\textcopyright}, % PDF Author
pdfsubject={}, % PDF Subject
pdfkeywords={}, % PDF Keywords
pdfcreator={pdfLaTeX}, % PDF Creator
pdfproducer={LaTeX with hyperref and ClassicThesis} % PDF producer
}

### 2014-10-09 R/clusterPrices.R

@@ -1,9 +1,16 @@
+### Unused preliminary clustering work
+library("dtw")
+library("gplots")
+library("pvclust")
+library("RColorBrewer")
+library("reshape2")
+library("snow")
+library("TSclust")
 source("R/functions.R")
 set.seed(2)
 
 ## cluster monthly price change profiles
 houses <- loadHouses()
-
 
 # sample 20 random postcodes for clustering
 # redraw london as clustered by price
@@ -17,9 +24,6 @@ ggplot(sub, aes(x=yearmon, y=Price/1e3, group=sector)) +
   facet_wrap(~sector, ncol=5, scales="free_y") +
   theme_sdc()
 
-
-head(houses)
-library("reshape2")
 hm <- dcast(houses, sector ~ yearmon, 
             fun.aggregate=median, value.var="Price")
 
@@ -27,9 +31,6 @@ hm <- dcast(houses, sector ~ yearmon,
 hm.mask <- dcast(houses, sector ~ yearmon,  value.var="Price")
 
 image(log10(as.matrix(hm[,-1])+1))
-
-library("gplots")
-library("RColorBrewer")
 
 num.cols=24
 
@@ -46,9 +47,6 @@ heatmap.2(log10(as.matrix(hm[complete.cases(hm),-1])),
           RowSideColors=heat.colors(max(rsc))[rsc],
           breaks=quantile(log10(houses$Price), seq(0.01, .97, length.out=(num.cols+1))))
 
-length(rsc)
-
-library("dtw")
 ts.dist <- function(x, ...)
   dist(x, method="DTW")
 
@@ -62,7 +60,7 @@ heatmap.2(mat, dendrogram="row", trace="none", Colv=NULL,
           distfun=ts.dist, hclustfun=hc, RowSideColors=rowcols,
           col=bl, labRow=F, labCol=row.lab, srtCol=45, breaks=brks)
 
-library("TSclust")
+
 sax.dist <- function(x, ...)
   # try dist on un-transformed values
   diss(as.matrix(hm[complete.cases(hm),-1]), "DWT", ...)
@@ -76,9 +74,6 @@ heatmap.2(mat, dendrogram="row", trace="none", Colv=NULL,
 
 dist.mat <- sax.dist(mat)
 
-
-library("pvclust")
-library("snow")
 cli <- makeCluster(6, type="MPI")
 clustered <- parPvclust(cli, t(mat), method.hclust="centroid",
                         method.dist="correlation", nboot=50)

### 2014-10-09 R/grading.R

@@ -112,3 +112,6 @@ hist(top.grades$vol, add=T, freq=F, col=rgb(.8,.5,.5,.2))
 
 df <- df[,c(1,2,4,3,5,8,6:7)]
 saveRDS(df, "rds/invest_grade.rds")
+
+
+head(df[df$sector=="SW18 4",])

### 2014-10-09 report/article_4.tex
some report and web tweaks

@@
 data with theoretical model outputs. These include South London
-suburbs of Upper Tooting and Morden, Walthamstow, as well as Brockley,
-whose train station within SE4 1 was linked with the London overground network in 2010.\\
+suburbs of Upper Tooting and Morden, as well as Walthamstow in East
+London, and Brockley,
+whose train station within SE4 1 was linked with the London Overground network in 2010.\\
 

### 2014-10-12 report/article_4.tex
report tweaks

@@
 were combined into an \emph{investment grade} for a given postcode
-area. Combined this set of analyses forms the basis of \logo, a
-startup aimed at ``democratising real estate investment'', offering
-individuals and businesses quantitative real estate insights for
+area. Together this set of analyses forms the basis of \logo, a
+startup aimed at \emph{democratising real estate investment}, offering
+individuals and businesses quantitative property area insights for
 data-driven investment decisions.
@@
 When considering a property in a given area, you may ask a realtor
-questions like: how do prices here compare to surrounding postcodes?
-How have prices increased over time? And how have prices changed over
-time? A real estate professional can give an opinion based on their
-own limited sample size, but with a large dataset we can address these
-questions quantitatively and visualise the results.
+questions like: how do prices here compare to surrounding postcodes, and
+how have they changed over time? A real estate professional can
+give an opinion based on their own limited sample size, but with a
+large dataset we can address these questions quantitatively and
+visualise the results.
 
@@
 Letting and sales sites may currently list some recent sales, but it's
-not currently possible to see a
+not currently possible to see a quantitative overview of property
+prices in a region, within the context of its sector, district and
+postcode.
 
@@
 
-\noindent quantitative overview of property
-prices in a region, within the context of its sector, district and
-postcode.
+%\noindent
 

### 2014-10-15 R/fractal_context.R
start tidyup

@@ -1,6 +1,5 @@
 library("directlabels")
 library("dplyr")
-library("ggplot2")
 library("RColorBrewer")
 library("reshape2")
 library("rgdal")
@@ -14,24 +13,12 @@ options(scipen=99)
 ## Sector (CB21 6):         2640       511
 ## Postocde (CB21 6XA):   305958         3
 
-# EC2Y 9AP - 228 house sales over 5 year period
-
-# SE1 6EG - 124
 top <- sort(table(houses$district), decreasing=T)
 head(top)
 
 # say you're looking at a house in: SW18 4HU 
 sw18 <- subset(houses, district == "SW18")
 sw184 <- subset(houses, sector == "SW18 4")
-
-sw18 <- subset(houses, district == "SW18")
-ggplot(sw18, aes(x=Trdate, y=Price)) +
-  geom_point(alpha=I(.2)) +
-  #scale_y_log10() +
-  geom_smooth(method="loess") +
-  coord_cartesian(ylim=c(quantile(sw18$Price, .05), 
-                         quantile(sw18$Price, .95)))
-
 
 modelArea <- function(data, dates, label){
   if (nrow(data) > 2e4){
@@ -61,15 +48,15 @@ all.res <- modelArea(houses, quarts, "South-East")
 
 qs <- rbind(london.res, sw.res, sw18.res, sw184.res, sw184hu.res, all.res)
 
+# Generate color palette w/ last two: grey shaded
 cols <- rev(brewer.pal(5, "Blues")[-1])
-# last two: grey shaded
 cols <- c(cols, "grey60", "grey80")
 
 qs$area <- factor(qs$area, 
             levels=c("SW18 4HU", "SW18 4", "SW18", "SW", "London", "South-East"))
 
-## static plots for now:
-pdf("plots/sw18.pdf", 5.5, 5.5)
+## Loess models for each area heirarchy (web)
+pdf("plots/PostcodeLevelsLoess.pdf", 5.5, 5.5)
 ggplot(qs, aes(x=time, y=fit/1000, col=area, fill=area)) +
   geom_ribbon(aes(ymin=(fit/1000) - (se.fit/1000),
                   ymax=(fit/1000) + (se.fit/1000)),
@@ -85,13 +72,10 @@ ggplot(qs, aes(x=time, y=fit/1000, col=area, fill=area)) +
[…]

### 2014-10-15 R/postcode_map.R

@@ -1,3 +1,5 @@
+######### Overview visualisation:: Web gifs #########
+library("ggmap")
 source("R/functions.R")
 
 houses <- loadHouses()
@@ -7,26 +9,13 @@ lat.max <- max(houses$Latitude)
 long.min <- min(houses$Longitude)
 long.max <- max(houses$Longitude)
 
-# base graphics data overview
-map("worldHires", ".", col="gray90", fill=TRUE, 
-    xlim=c(long.min, long.max), ylim=c(lat.min, lat.max))
-points(p$longitude, p$latitude, pch=".")
-
-
-
-library("ggmap")
 map <- get_map(location="London", zoom=8, maptype="toner", color="bw", source="stamen")
 p1 <- ggmap(map, extent="device", darken=.6)
-p1 + geom_point(data=houses[sample(1:nrow(houses), 1e5),], 
-                aes(x=Longitude, y=Latitude, alpha=log10(Price)), col="white") +
-  scale_alpha_continuous(range=c(.01, .2)) +
-  labs(x="", y="") + theme(legend.position="none") 
 
 # filename ordering for (e.g.) imagemagick (yes, 51 not 52)
 getfn <- function(n)
   ifelse(n > 51, paste0("zz", letters[n %% 26]), 
          ifelse(n > 26, paste0("z", letters[n %% 26]), letters[n]))
-
 
 for( ym in 1:65 ){
   # x=1, y=50.5

### 2014-10-15 R/report_viz.R

@@ -1,3 +1,4 @@
+######### Overview visualisation:: Report intro #########
 library("ggmap")
 source("R/functions.R")
 set.seed(42)

### 2014-10-15 R/functions.R
more tidyup

@@ -8,7 +8,7 @@ library("rgeos")
 library("zoo")
 
 loadHouses <- function(){
-  h <- read.csv("houseprices/london2009-2014-house-prices//Houseprice_2009_100km_London.csv", 
+  h <- read.csv("houseprices/Houseprice_2009_100km_London.csv", 
                 stringsAsFactors=F)
   
   # remove underscores, convert to number

### 2014-10-16 report/article_4.tex
start tidyup
latex report expansion

@@
 
-% \begin{figure}[h!]
-% \begin{center}
-% \includegraphics[width=.1\textwidth]{Figures/logo.png} \\
-% \includegraphics[scale=.12]{Figures/logotext.png} \\
-% \caption*{ {\sf Democratising real estate investment}}
-% \end{center}
-% \end{figure}
+\clearpage
+
+\section*{Summary}
+
+In this report I've given a brief overview my attempts to elute maximal
+insight from a single large dataset through the data science techniques of
+visualisation and quantitative modelling. Together they form a
+cohesive narrative of investigative analysis into a single exemplar
+postcode, and discovery of the potential investment returns
+therein.
+
+These analyses have to potential to generate valuable insights for a range of
+customer segments:
+\begin{itemize}
+
+\item Online sales and letting agents would benefit from
+plug-in analytics provided by \logo \hspace{.1em} --- offering their
+users advanced insights into spatiotemporal property statistics.
+
+\item A web and mobile app to make a set of analyses widely-available
+  to amateur property investors, new homebuyers and landlords.
+
+\item A consultancy specialising in real estate investment to partner
+  with REITs and other specialist investment funds, as well as charitable trusts.
+\end{itemize}
 
@@
 
+\vspace{2em}
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=.15\textwidth]{Figures/logo.png} \\
+\includegraphics[scale=.12]{Figures/logotext.png} \\
+\caption*{ {\sf Democratising real estate investment}}
+\end{center}
+\end{figure}
+
 \end{document}

### 2014-10-20 report/article_4.tex
new report figure; cut-down

@@
 
-A full report with additional visualisations is online at {\bf
+A full report with additional interactive visualisations is online at {\bf
   \leavevmode\href{http://blm.io/datarea}{blm.io/datarea} }and scripts to
@@
 
+\vspace{-.5em}
 \subsection*{Fractal context}
@@
 
+\vspace{-.5em}
 \subsection*{ARIMA modelling}
@@
 
-\vspace{-1em}
+\vspace{-2em}
 \subsection*{Investment grading}
@@
 
-As an example, Figure 3 shows the top 5 best-ranked postcode sectors
+As an example, Figure 3 shows the top 5 postcode sectors
 in which to invest, according to their 12-month price growth forecast
 and historical annualised volatility --- thus combining the empirical
-data with theoretical model outputs. These include South London
-suburbs of Upper Tooting and Morden, as well as Walthamstow in East
-London, and Brockley,
+data with theoretical model outputs. These include Brockley,
 whose train station within SE4 1 was linked with the London Overground network in 2010.\\
@@
 
-The rankings used in Figure 3 made use of the $drift$ parameter of
-ARIMA models, combined with statistical volatility over the five year
-house prices dataset. Both parameters were converted to quantiles and
+\setlength{\intextsep}{0em}
+\begin{wrapfigure}{R}{.45\textwidth}
+%\vspace{-1em}
+\centering
+\includegraphics[width=.44\textwidth]{Figures/grading.png}
+\vspace{-1em}
+\caption{ Quantile transformation for investment grading
+  (AAA rated sectors in \emph{navy}).}
+\vspace{.5em}
+\end{wrapfigure}
+
+Growth forecasts and volatility were converted to quantiles and
 weighted equally; those sectors with the highest projected growth
-and least historical volatility receive the highest investment grade
+and least historical volatility receive the best investment grade
 (on a scale of AAA, AA, A, BBB \ldots C).\textsuperscript{$\star$}
 
-Returning to our original example, we can conclude that SW18 4HU is a
-promising choice of area for property investment. It's approximately
-median-priced for the sector, which is the cheapest of SW18 (in turn
-part of the desirable SW London area; Figure 1). It has good growth
-projections of XX in the next 12 months (Figure 2), and when combined
-with volatility and compared with all $\approx2600$ sectors in the
-dataset, receives the top AAA investment grade, placing it in the
-90\textsuperscript{th} percentile for predicted price growth and low
-historical volatility.
+Returning to our original example, SW18 4HU is a promising choice of
+area for property investment. It's median-priced for the sector, which
+is the cheapest of SW18 (in turn part of the desirable SW London area;
+Figure 1). It has good growth projections of over £28,000 in the next
+year (Figure 2), and receives the top AAA investment grade, being
+in the 90\textsuperscript{th} percentile for predicted price growth
+and low historical volatility in the dataset.
 

### 2014-10-20 R/interactive.R

@@ -1,0 +1,38 @@
+#require(devtools)
+#install_github('ramnathv/rCharts')
+library("rCharts")
+df <- readRDS("rds//invest_grade2.rds")
+
+dfi <- df[df$growth >0,]
+dfi$growth.q <- ecdf(dfi$growth)(dfi$growth)
+
+colnames(dfi)
+colnames(dfi)[6] <- "Grade"
+colnames(dfi)[4] <- "Growth.quantile"
+colnames(dfi)[5] <- "Volatility.quantile"
+
+dfi[,2:5] <- apply(dfi[,2:5], 2, round, digits=2)
+dfi$rank <- NULL
+dfi$sum <- NULL
+
+d <- dPlot(Growth.quantile ~ Volatility.quantile, type="bubble", data=dfi,
+      col="Grade", height=450, width=450, groups=c("sector", "Grade"))
+d$xAxis(type = "addMeasureAxis")
+d$yAxis(type = "addMeasureAxis")
+d
+
+d$publish()
+
+
+colnames(dfi)[2:3] <- c("Monthly.growth", "Historical.volatility")
+dfii <- dfi[,c(1:6)]
+d2 <- dPlot(Monthly.growth ~ Historical.volatility, type="bubble", data=dfii,
+           col="Grade", height=450, width=450, groups=c("sector", "Grade"))
+d2$xAxis(type = "addMeasureAxis")
+#d2$xAxis(type  = "addLogAxis", overrideMin=.2, overrideMax=5)
+d2$yAxis(type = "addMeasureAxis")
+#d2$yAxis(type = "addLogAxis", overrideMin=10)
+d2
+d2$publish()
+
+

### 2014-10-20 R/investment_grade.R

@@ -52,12 +52,6 @@ df$growth <- pvec(as.vector(df$sector), growth, mc.cores=8)
 ## convert raw growth vals to quantiles
 df$growth.q <- ecdf(df$growth)(df$growth)
 
-hist(df$growth.q)
-
-tc <- houses[houses$sector == "NW9 6",]
-plot(tc$Trdate, tc$Price)
-plot(tc$Trdate, fitted(auto.arima(ts(tc$Price, as.numeric(tc$Trdate)))))
-
 volatility <- function(sector){
   data <- houses[houses$sector == sector,]
   data <- group_by(data, Month) %>% summarise(median=median(Price))
@@ -99,19 +93,8 @@ svg("plots/top5_investments.svg", 8, 7)
 do.call(grid.arrange, l)
 dev.off()
 
-head(top.grades)
-hist(rs <- rowSums(df[,c("growth.q", "vol.q")]))
-which.max(rs)
-df[1168,]
-
-plot(density(df$growth))
-lines(density(top.grades$growth))
-
-hist(df$vol, 50, freq=F, ylim=c(0,1.5), col=rgb(.5,.5,.8,.2))
-hist(top.grades$vol, add=T, freq=F, col=rgb(.8,.5,.5,.2))
 
 df <- df[,c(1,2,4,3,5,8,6:7)]
-saveRDS(df, "rds/invest_grade.rds")
+saveRDS(df, "rds/invest_grade2.rds")
 
 
-head(df[df$sector=="SW18 4",])

### 2014-10-20 build.sh

@@ -1,0 +1,2 @@
+echo rm -R ~/other/blmoore.github.io/datarea/*
+echo cp -R writeup/* ~/other/blmoore.github.io/datarea/.

### 2014-10-26 gmap/map.js
gmap

@@ -1,0 +1,2023 @@
+var hmdata = [
+{location: new google.maps.LatLng(51.15, 0.27), weight:2.2},
+{location: new google.maps.LatLng(50.84, 0.43), weight:1.7},
+{location: new google.maps.LatLng(51.43, -0.26), weight:2.3},
+{location: new google.maps.LatLng(51.48, -0.44), weight:1.9},
+{location: new google.maps.LatLng(51.28, -0.96), weight:1.8},
+{location: new google.maps.LatLng(51.31, -0.24), weight:2.2},
+{location: new google.maps.LatLng(51.46, -1.03), weight:1.8},
+{location: new google.maps.LatLng(51.66, -0.9), weight:1.7},
+{location: new google.maps.LatLng(52.12, -0.46), weight:1.5},
+{location: new google.maps.LatLng(51.41, -1.26), weight:1.7},
+{location: new google.maps.LatLng(51.41, -0.29), weight:1.9},
+{location: new google.maps.LatLng(51.23, -0.17), weight:1.8},
+{location: new google.maps.LatLng(51.57, 0.55), weight:1.8},
+{location: new google.maps.LatLng(51.4, -1.33), weight:1.3},
+{location: new google.maps.LatLng(51.43, -0.53), weight:1.7},
+{location: new google.maps.LatLng(50.84, -0.21), weight:1.7},
+{location: new google.maps.LatLng(51.64, -0.18), weight:2.1},
+{location: new google.maps.LatLng(51.32, -0.73), weight:1.8},
+{location: new google.maps.LatLng(51.25, -0.66), weight:2.3},
+{location: new google.maps.LatLng(51.41, 0.01), weight:1.7},
+{location: new google.maps.LatLng(51.57, -0.35), weight:1.6},
+{location: new google.maps.LatLng(51.5, -0.08), weight:1.9},
+{location: new google.maps.LatLng(51.52, -0.09), weight:2.2},
+{location: new google.maps.LatLng(51.04, 0.5), weight:1.7},
+{location: new google.maps.LatLng(51.47, 0.01), weight:2.5},
+{location: new google.maps.LatLng(51.8, 0.63), weight:1.5},
+{location: new google.maps.LatLng(51.52, -0.6), weight:1.7},
+{location: new google.maps.LatLng(51.58, 0.41), weight:1.3},
+{location: new google.maps.LatLng(51.55, -0.11), weight:2.1},
[…]

### 2014-10-26 wip/gmap.R

@@ -1,0 +1,43 @@
+source("R/functions.R")
+
+houses <- loadHouses()
+
+# grad 10 points for heatmap
+ll <- houses[sample(1:nrow(houses), 2000),c("Latitude", "Longitude", "Price")]
+ll[,1:2] <- apply(ll[,1:2], 2, round, digits=2)
+
+weights <- round(log10(ll$Price), digits=1)
+weights <- weights - min(weights) + 1
+
+raw <- 
+  paste0("{location: new google.maps.LatLng(", 
+         ll[,1], ", ", ll[,2], "), weight:", weights, "},")
+
+
+clip <- pipe("pbcopy", "w")
+write(raw, file=clip)
+close(clip)
+
+set.seed(42)
+options(scipen=99)
+h <- houses[,c("Latitude", "Longitude", "Price", "Trdate")]
+colnames(h)[4] <- "Date"
+write.csv(h, "gmap/fusiondata_full.csv", row.names=F)
+
+## sector KML
+kml <- read.csv("gmap/UK postcode districts.csv")
+length(unique(houses$sector))
+kml <- kml[kml[,1] %in% unique(houses$district),]
+
+library("dplyr")
+k <- group_by(houses, district) %>% 
+  filter(district %in% kml[,1]) %>%
+  summarise(median=median(Price), number=n(),
+            newbuilds=length(Newbuild[Newbuild == "Y"]),
+            min=min(Price), max=max(Price))
+
+# match order (AL10 195000)
+k <- k[match( as.character(kml[,1]), k$district ),]
+k$kml <- paste0("\"", kml[,2], "\"")
+k$median <- round(k$median, -3)
+write.csv(k, "gmap/fusion_kml.csv", row.names=F, quote=F)

### 2014-10-27 wip/gmap.R
new report figure

@@ -34,7 +34,8 @@ k <- group_by(houses, district) %>%
   filter(district %in% kml[,1]) %>%
   summarise(median=median(Price), number=n(),
             newbuilds=length(Newbuild[Newbuild == "Y"]),
-            min=min(Price), max=max(Price))
+            min=min(Price), max=max(Price),
+            mean=round(mean(Price), -3))
 
 # match order (AL10 195000)
 k <- k[match( as.character(kml[,1]), k$district ),]

### 2014-11-03 report/article_4.tex
gmap
new report figure
rm unneeded files; add some comments

@@
 
-I analysed the {\bf London house prices} dataset is several innovative
+I analysed the {\bf London house prices} dataset in several innovative
 ways. Firstly \emph{fractal context} shows how house sales within an
@@
 startup aimed at \emph{democratising real estate investment}, offering
-individuals and businesses quantitative property area insights for
+individuals and businesses quantitative property area insights to enable
 data-driven investment decisions.
@@
 \vspace{-1em}
-\section*{Analyses}
+\section*{Section A: Analyses}
 
@@
 \begin{wrapfigure}{R}{.45\textwidth}
-%\vspace{-1em}
+\vspace{.5em}
 \centering
-\includegraphics[width=.44\textwidth]{Figures/grading.png}
+\includegraphics[width=.44\textwidth]{inkscape/grading_2.png}
 \vspace{-1em}
@@
 \clearpage
-\section*{Value generation}
+\section*{Section B: Value generation}
 
@@
 
+\begin{figure}[h]
+\vspace{1.5em}
+\centering
+\includegraphics[width=.85\textwidth]{Figures/gmapScreenshot.png}
+\vspace{.5em}
+\caption{ A basic embeddable web app built with the Google Maps API
+  (viewable at \href{http://blm.io/datarea}{blm.io/datarea}). }
+\end{figure}
+
 \subsection*{2. Direct-to-consumer demo application}
+
 \logo \hspace{.1em} would release both a {\bf web service} and associated
@@
 
-\clearpage
-
 \section*{Summary}
 
-In this report I've given a brief overview my attempts to elute maximal
-insight from a single large dataset through the data science techniques of
-visualisation and quantitative modelling. Together they form a
-cohesive narrative of investigative analysis into a single exemplar
-postcode, and discovery of the potential investment returns
-therein.
+\begin{wrapfigure}[10]{r}{.3\textwidth}
+\vspace{2em}
+\centering
+\includegraphics[width=.2\textwidth]{Figures/logo.png}
+\includegraphics[width=.2\textwidth]{Figures/logotext.png}
+%\caption*{ {\sf Democratising real estate investment}}
+\caption{ Logo design. }
+%\vspace{-4em}
+\end{wrapfigure}
 
-These analyses have to potential to generate valuable insights for a range of
+In this report I've given a brief overview my attempts to elute
+maximal insight from a single large dataset through visualisations and
+quantitative modelling. Together these form a cohesive narrative of
+investigative analysis into a single exemplar postcode, and discovery
+of the potential investment returns therein.
+
+These analyses have the potential to generate valuable insights for a range of
 customer segments:
@@
 \item A consultancy specialising in real estate investment to partner
-  with REITs and other specialist investment funds, as well as charitable trusts.
+  with REITs and other specialist investment funds, as well as
+  charitable trusts.
 \end{itemize}
@@
 
-\vspace{2em}
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=.15\textwidth]{Figures/logo.png} \\
-\includegraphics[scale=.12]{Figures/logotext.png} \\
-\caption*{ {\sf Democratising real estate investment}}
-\end{center}
-\end{figure}
+%\vspace{2em}
+% \begin{figure}[h!]
+% \begin{center}
+% \includegraphics[width=.15\textwidth]{Figures/logo.png} \\
+% \includegraphics[scale=.11]{Figures/logotext.png} \\
+% \caption*{ {\sf Democratising real estate investment}}
+% \end{center}
+% \end{figure}
 

### 2014-11-03 R/arima_model.R

@@ -1,4 +1,10 @@
-## ARIMA forecasting
+### Time series modelling and forecasting ###
+# Performs some data munging and diagnostic #
+# tests to help decide the best modelling   #
+# approach. Then fits a regularised ARIMA   #
+# model to the example sector ("SW18 4"),   #
+# plotting the forecast (report: fig. 2).   #
+#############################################
 library("dplyr")
 library("forecast")
 library("ggplot2")
@@ -69,10 +75,8 @@ afit <- data.frame(Month=sdf$Month, median=as.vector(fitted(aa)))
 
 pdf$type <- factor(pdf$type, levels=c("measured", "forecast"))
 
-## hacky ggplot abuse
 pdf("plots/forecast_v2.pdf", 5, 3)
 ggplot(pdf, aes(x=Month, y=median/1e3)) +
-  #scale_color_manual() +
   # forecase bounds
   geom_ribbon(data=vars, inherit.aes=F,
               aes(x=Month, ymin=l95/1e3, ymax=h95/1e3),
@@ -91,7 +95,6 @@ ggplot(pdf, aes(x=Month, y=median/1e3)) +
   annotate("text", x=as.Date("2010-01-01"), size=8,
            y=550, label="SW18 4", col=I("grey40"))
 dev.off()
-
 
 plotForecast <- function(out){
   pdf <- out$pdf
@@ -122,14 +125,12 @@ plotForecast <- function(out){
 }
 
 modelHPs <- function(sector){
-  #sector="W6 7"
   data <- houses[houses$sector == sector,]
   # aggregate by monht
   data <- group_by(data, Month) %>% summarise(median=median(Price))
   data.ts <- ts(data$median, as.numeric(as.yearmon(data$Month)))
   arima <- auto.arima(data.ts)
   # inspect fit
-  # summary(arima)
   f <- forecast(arima, h=12)
   fore <- as.data.frame(f)
   
@@ -157,7 +158,6 @@ modelHPs <- function(sector){
 # random selection of sectors:
 set.seed(42)
 sectors <- c("SW18 4", sample(houses$sector, 15))
-
 l <- lapply(sectors, modelHPs)
 
 svg("plots/grid_forecasts_v2.svg", 16, 10)

### 2014-11-03 R/datablog.R

@@ -1,27 +1,0 @@
-boroughs <- read.csv("datablog_data/London boroughs in numbers.csv")
-
-# relate boroughs to postcodes via london_postcodes.csv from: http://www.doogal.co.uk/london_postcodes.php
-b2pc <- read.csv("London_postcodes.csv")
-head(b2pc)
-boroughs$BOROUGH
-
-# b2pc$District \approx boroughs
-
-for( b in unique(boroughs$BOROUGH)) {
-  cat(b, "\n")
-  m <- unique(b2pc[agrep(b, b2pc$District),]$District)
-  cat(m, b, "\n") 
-}
-
-b2pc$District[agrep(boroughs$BOROUGH[1], b2pc$District)]
-
-
-earnings <- read.csv("datablog_data/Figure 3.15, Trends in median full-time gross weekly earnings of employees by gender, 1966-2008.xls.csv")
-
-london.wards <- readShapePoly("~/TD/london_wards2013/london_wards2013.shp"
-                              , proj4string=CRS(projString))
-wards.count <- nrow(london.wards@data)
-# assign id for each lsoa
-
-london.wards@data$id <- 1:wards.count
-wards.fort <- fortify(london.wards, region='id')

### 2014-11-03 R/fractal_context.R

@@ -1,3 +1,19 @@
+# ##########################################################
+# ------------------- Fractal context -------------------- #
+#                                                          #
+# Draws individual plots which together give a "fractal    #
+# context" of a given postcode area. This involves showing #
+# price distributions throughout the postcode hierarchy:   #
+#                                                          #
+#    FC0 :: violin plots of London areas                   #
+#    FC1 :: a colour-coded map of SW London                #
+#    FC2 :: violin densitites with overlaid individual     #
+#           house sales for each sector                    #
+#    FC3 :: a kernel-density estimate of the distribution  #
+#           of house sale prices within SW18 4             #
+#                                                          #
+# These vector graphs were then arranged using Inkscape.   #
+############################################################
 library("directlabels")
 library("dplyr")
 library("RColorBrewer")

### 2014-11-03 R/functions.R

@@ -62,7 +62,6 @@ theme_sdc_null <- function(base_size=12, base_family="sans") {
     axis.line.y = element_line(),
     legend.background = element_rect(colour = "grey95", fill="grey95"),
     legend.margin = grid::unit(0.2, "cm"),
-    # legend.key = element_rect(),
     legend.key = element_blank(),
     legend.key.size = grid::unit(1.2, "lines"),
     legend.key.height = NULL,
@@ -76,9 +75,7 @@ theme_sdc_null <- function(base_size=12, base_family="sans") {
     legend.justification = "center",
     legend.box = NULL,
     ## Must have colour=NA or covers the plot
-    # panel.background = element_rect(),
     panel.background = element_rect(fill = "transparent", colour = NA),
-    # panel.border = element_rect(fill=NA),
     panel.border = element_blank(), 
     panel.margin = grid::unit(0.25, "lines"),
     panel.grid = element_line(),
@@ -86,9 +83,6 @@ theme_sdc_null <- function(base_size=12, base_family="sans") {
     panel.grid.major.x = element_line(),
     panel.grid.major.y = element_line(),
     # removed:
-    #panel.grid.minor = element_line(),
-    #panel.grid.minor.x = element_line(size=.1),
-    #panel.grid.minor.y = element_line(size=.1),
     panel.grid.minor = element_blank(),
     panel.grid.minor.x = element_blank(),
     panel.grid.minor.y = element_blank(),
@@ -117,24 +111,6 @@ theme_sdc_sizes <- function() {
   )
 }
 
-#' sdc custom ggplot2 theme
-#'
-#' Custom ggplot2 theme for #summerdatachallenge.
-#' Reasonably minimalist but retains grid lines which might
-#' be better off stripped depending on data.
-#' 
-#' @export
-#' @examples
-#' library("ggplot2")
-#' library("grid")
-#' 
-#' # default theme
-#' ggplot(data.frame(x=1:100, y=rnorm(100)), aes(x, y)) + 
-#'   geom_point()
-#'  
-#' #sdc
-#' ggplot(data.frame(x=1:100, y=rnorm(100)), aes(x, y)) + 
-#'   geom_point() + theme_sdc()
 theme_sdc <- function(base_size=12, base_family="", use_sizes=TRUE) {
   thm <- theme_sdc_null(base_size=base_size, base_family=base_family)
   if (use_sizes) {
[…]

### 2014-11-03 R/interactive.R

@@ -1,7 +1,13 @@
+######## Draw interactive js plots ########
+# Generates interactive javascript plots  #
+# for web write-up. Base dimple.js charts #
+# published here, then fixed-up manually. #
+###########################################
+
 #require(devtools)
 #install_github('ramnathv/rCharts')
 library("rCharts")
-df <- readRDS("rds//invest_grade2.rds")
+df <- readRDS("rds//invest_grade.rds")
 
 dfi <- df[df$growth >0,]
 dfi$growth.q <- ecdf(dfi$growth)(dfi$growth)
@@ -15,6 +21,7 @@ dfi[,2:5] <- apply(dfi[,2:5], 2, round, digits=2)
 dfi$rank <- NULL
 dfi$sum <- NULL
 
+## Raw data
 d <- dPlot(Growth.quantile ~ Volatility.quantile, type="bubble", data=dfi,
       col="Grade", height=450, width=450, groups=c("sector", "Grade"))
 d$xAxis(type = "addMeasureAxis")
@@ -23,16 +30,15 @@ d
 
 d$publish()
 
-
+## Quantiles
 colnames(dfi)[2:3] <- c("Monthly.growth", "Historical.volatility")
 dfii <- dfi[,c(1:6)]
 d2 <- dPlot(Monthly.growth ~ Historical.volatility, type="bubble", data=dfii,
            col="Grade", height=450, width=450, groups=c("sector", "Grade"))
 d2$xAxis(type = "addMeasureAxis")
-#d2$xAxis(type  = "addLogAxis", overrideMin=.2, overrideMax=5)
 d2$yAxis(type = "addMeasureAxis")
-#d2$yAxis(type = "addLogAxis", overrideMin=10)
 d2
+
 d2$publish()
 
 

### 2014-11-03 R/investment_grade.R

@@ -25,12 +25,7 @@ growth <- function(sector){
   data <- group_by(data, Month) %>% summarise(median=median(Price))
   data.ts <- ts(data$median, as.numeric(as.yearmon(data$Month)))
   arima <- auto.arima(data.ts, allowdrift=T)
-#  f <- as.data.frame(forecast(arima, h=12))
-  #s.e <- fore$`Point Forecast`[c(1, 12)]
-  
-  # percentage change over forecast
-  #100*((s.e[2] - s.e[1]) / s.e[1])
- # message(coef(arima))
+
   if("drift" %in% names(coef(arima))){
     return(coef(arima)[["drift"]])
   } else {
@@ -95,6 +90,6 @@ dev.off()
 
 
 df <- df[,c(1,2,4,3,5,8,6:7)]
-saveRDS(df, "rds/invest_grade2.rds")
+saveRDS(df, "rds/invest_grade.rds")
 
 

### 2014-11-03 R/theatre.R

@@ -1,10 +1,0 @@
-sales <- read.csv("theatre//london2012-theatre-v2//Sales Data.csv",
-                  colClasses=c(rep("character", 3),
-                               "integer", "numeric", 
-                               rep("integer", 2)))
-str(sales)
-
-sales$sale_date <- as.Date(sales$sale_date)
-sales$performance_date <- as.Date(sales$performance_date)
-
-hist(sales$total_tickes)

### 2014-11-03 build.sh

@@ -1,2 +1,0 @@
-echo rm -R ~/other/blmoore.github.io/datarea/*
-echo cp -R writeup/* ~/other/blmoore.github.io/datarea/.

### 2014-11-03 gmap/map.js

@@ -1,2023 +1,0 @@
-var hmdata = [
-{location: new google.maps.LatLng(51.15, 0.27), weight:2.2},
-{location: new google.maps.LatLng(50.84, 0.43), weight:1.7},
-{location: new google.maps.LatLng(51.43, -0.26), weight:2.3},
-{location: new google.maps.LatLng(51.48, -0.44), weight:1.9},
-{location: new google.maps.LatLng(51.28, -0.96), weight:1.8},
-{location: new google.maps.LatLng(51.31, -0.24), weight:2.2},
-{location: new google.maps.LatLng(51.46, -1.03), weight:1.8},
-{location: new google.maps.LatLng(51.66, -0.9), weight:1.7},
-{location: new google.maps.LatLng(52.12, -0.46), weight:1.5},
-{location: new google.maps.LatLng(51.41, -1.26), weight:1.7},
-{location: new google.maps.LatLng(51.41, -0.29), weight:1.9},
-{location: new google.maps.LatLng(51.23, -0.17), weight:1.8},
-{location: new google.maps.LatLng(51.57, 0.55), weight:1.8},
-{location: new google.maps.LatLng(51.4, -1.33), weight:1.3},
-{location: new google.maps.LatLng(51.43, -0.53), weight:1.7},
-{location: new google.maps.LatLng(50.84, -0.21), weight:1.7},
-{location: new google.maps.LatLng(51.64, -0.18), weight:2.1},
-{location: new google.maps.LatLng(51.32, -0.73), weight:1.8},
-{location: new google.maps.LatLng(51.25, -0.66), weight:2.3},
-{location: new google.maps.LatLng(51.41, 0.01), weight:1.7},
-{location: new google.maps.LatLng(51.57, -0.35), weight:1.6},
-{location: new google.maps.LatLng(51.5, -0.08), weight:1.9},
-{location: new google.maps.LatLng(51.52, -0.09), weight:2.2},
-{location: new google.maps.LatLng(51.04, 0.5), weight:1.7},
-{location: new google.maps.LatLng(51.47, 0.01), weight:2.5},
-{location: new google.maps.LatLng(51.8, 0.63), weight:1.5},
-{location: new google.maps.LatLng(51.52, -0.6), weight:1.7},
-{location: new google.maps.LatLng(51.58, 0.41), weight:1.3},
-{location: new google.maps.LatLng(51.55, -0.11), weight:2.1},
[…]

### 2014-11-03 wip/gmap.R

@@ -1,35 +1,17 @@
+######### Google maps API web app #########
+# Generates input csv for a fusion table  #
+# to be used with google maps API.        #
+###########################################
+library("dplyr")
 source("R/functions.R")
 
 houses <- loadHouses()
-
-# grad 10 points for heatmap
-ll <- houses[sample(1:nrow(houses), 2000),c("Latitude", "Longitude", "Price")]
-ll[,1:2] <- apply(ll[,1:2], 2, round, digits=2)
-
-weights <- round(log10(ll$Price), digits=1)
-weights <- weights - min(weights) + 1
-
-raw <- 
-  paste0("{location: new google.maps.LatLng(", 
-         ll[,1], ", ", ll[,2], "), weight:", weights, "},")
-
-
-clip <- pipe("pbcopy", "w")
-write(raw, file=clip)
-close(clip)
-
-set.seed(42)
-options(scipen=99)
-h <- houses[,c("Latitude", "Longitude", "Price", "Trdate")]
-colnames(h)[4] <- "Date"
-write.csv(h, "gmap/fusiondata_full.csv", row.names=F)
 
 ## sector KML
 kml <- read.csv("gmap/UK postcode districts.csv")
 length(unique(houses$sector))
 kml <- kml[kml[,1] %in% unique(houses$district),]
 
-library("dplyr")
 k <- group_by(houses, district) %>% 
   filter(district %in% kml[,1]) %>%
   summarise(median=median(Price), number=n(),
@@ -37,7 +19,7 @@ k <- group_by(houses, district) %>%
             min=min(Price), max=max(Price),
             mean=round(mean(Price), -3))
 
-# match order (AL10 195000)
+# match order of k and kml
 k <- k[match( as.character(kml[,1]), k$district ),]
 k$kml <- paste0("\"", kml[,2], "\"")
 k$median <- round(k$median, -3)

### 2014-11-03 R/arima_model.R
tidy; rm old files

@@ -160,7 +160,7 @@ set.seed(42)
 sectors <- c("SW18 4", sample(houses$sector, 15))
 l <- lapply(sectors, modelHPs)
 
-svg("plots/grid_forecasts_v2.svg", 16, 10)
+svg("plots/grid_forecasts.svg", 16, 10)
 do.call(grid.arrange, l)
 dev.off()
 

### 2014-11-07 report/sdc_report.tex
tidy; rm old files
tidy, rm big files

--- 2014-11-03 report/article_4.tex
@@
 
-% Make header in inkscape?
 \begin{figure}[width=\textwidth]
@@
 
-I analysed the {\bf London house prices} dataset in several innovative
-ways. Firstly \emph{fractal context} shows how house sales within an
-arbitrary area relate to those if neighbouring districs, up and down
-the postcode hierarchy. Secondly, I fit \emph{ARIMA models} to
-forecast an area's price trend into the future, with an eye to
-suggesting profitable investments. Finally, these and other metrics
+I analysed the {\bf London house prices} dataset in several creative
+ways. Firstly, the \emph{fractal context} of an area visualises how house
+prices relate to those of neighbouring areas at different scales up
+and down the postcode hierarchy. Secondly, I fit \emph{ARIMA models}
+to forecast an area's price trend into the future, with an eye to
+suggesting profitable investments. Finally, growth and other metrics
 were combined into an \emph{investment grade} for a given postcode
@@
 startup aimed at \emph{democratising real estate investment}, offering
-individuals and businesses quantitative property area insights to enable
-data-driven investment decisions.
+individuals and businesses quantitative property area insights to
+enable data-driven investment decisions.
 
-A full report with additional interactive visualisations is online at {\bf
-  \leavevmode\href{http://blm.io/datarea}{blm.io/datarea} }and scripts to
-reproduce all analyses are available from {\bf
-  \leavevmode\href{http://github.com/blmoore/summerdatachallenge}{github.com/blmoore/summerdatachallenge}}.
+An enhanced report with interactive visualisations is online at 
+{\bf \leavevmode\href{http://blm.io/datarea}{blm.io/datarea} } 
+and scripts to reproduce all analyses are available from 
+{\bf
+  \leavevmode\href{http://github.com/blmoore/summerdatachallenge}{github.com/blmoore/summerdatachallenge}
+}.
 
@@
 
-{\let\thefootnote\relax\footnotetext{\textsuperscript{1} \textit{MRC
-      Human Genetics Unit, University of Edinburgh, Scotland, United Kingdom}}}
+%{\let\thefootnote\relax\footnotetext{\textsuperscript{1} \textit{MRC
+ %     Human Genetics Unit, University of Edinburgh, Scotland, United Kingdom}}}
 
@@
 
-When considering a property in a given area, you may ask a realtor
-questions like: how do prices here compare to surrounding postcodes, and
-how have they changed over time? A real estate professional can
-give an opinion based on their own limited sample size, but with a
-large dataset we can address these questions quantitatively and
-visualise the results.
+When considering a property in a given area, you may want to ask
+questions like: how do prices here compare to the surrounding areas?
+How have they changed over time? Where will they be in a year's time?
+A real estate professional can give an opinion based on their own
+limited sample size, but with a large dataset we can address these
+questions through statistical analysis and visualise the results.
 
@@
 
-Letting and sales sites may currently list some recent sales, but it's
-not currently possible to see a quantitative overview of property
-prices in a region, within the context of its sector, district and
-postcode.
+Letting and sales sites sometimes list some recent sales in a given
+area, but it's not currently possible to get a quantitative overview
+of property prices in a region, within the context of its sector,
+district and postcode.
 
@@
 \begin{wrapfigure}[16]{l}{.45\textwidth}
-\vspace{.5em}
+%\vspace{.5em}
 \centering
@@
 These questions can be answered with a clear and intuitive data
-visualisation of price distributions in the postcode hierarchy --- we
-call this the fractal context of a property price (Figure 1).
+visualisation of price distributions in the postcode hierarchy --- what
+I've called the fractal context of a property price (Figure 1).
 
-A specific postcode or area can be marked within its distribution,
-giving a clear view of a properties relative pricing which cannot be
-gleaned just by browsing recent sales. The example view (Figure 1)
-combines ranked violin density plots and a geo-heatmap for a clear,
-intuitive view of a locations price context.
+Any chosen area can be seen in the context of its surroundings, from
+full postcodes to much wider areas, clearly showing how the pricing of
+any given property relates to those nearby.  The example graphic for
+postcode SW18 4HU (Figure 1) combines ranked violin density plots and
+a geo-heatmap for a clear, intuitive view of a location's price
+context.
 
@@
 
-Given a time series of housing prices, it's of interest to investors
-to speculate on potential investment returns for a given property
-area. This can be done through time series analysis using the
-autoregressive integrated moving average (ARIMA) model (Figure 2). The
-model fitting procedure included regularisation that allowed a $drift$
-term to model non-stationary trends, as well as coefficients capturing
-periodicity or seasonal effects. Potential investment returns could be
-optimised by selecting areas with
-maximal growth projections. \\
+Property investors need a method of evaluating the potential investment
+returns of a given property area. This can be achieved through time
+series modelling with autoregressive integrated moving average (ARIMA)
+models (Figure 2). For this application regularised models were fit
+which allowed a $drift$ term to capture non-stationary trends
+(i.e. allowing for gain, loss or neither) as well as coefficients
+reflecting periodicity or seasonal effects. Hence potential investment
+returns could then be optimised by selecting areas with
+maximal growth projections.
 
@@
 
-Combining growth forecasts and other derived metrics from this
-dataset allow an approximate investment grading, relative to other
-postcodes within the area covered by this dataset. This intuitive
-output metric can help prioritise areas ripe for property investment.
+Combining growth forecasts and with an additional measure of
+variability allows an approximate investment grading of each sector,
+relative to other postcodes within the area covered by this
+dataset. This intuitive output metric can help prioritise areas ripe
+for property investment and combines the theoretical model projections
+with an empirical metric of past stability.
 
-As an example, Figure 3 shows the top 5 postcode sectors
-in which to invest, according to their 12-month price growth forecast
-and historical annualised volatility --- thus combining the empirical
-data with theoretical model outputs. These include Brockley,
+As an example, Figure 3 shows the top 5 postcode sectors in which to
+invest, according to their 12-month price growth forecast and
+historical annualised volatility. These include Brockley,
 whose train station within SE4 1 was linked with the London Overground network in 2010.\\
@@
 Growth forecasts and volatility were converted to quantiles and
-weighted equally; those sectors with the highest projected growth
+equally weighted (Figure 4); those sectors with the highest projected growth
 and least historical volatility receive the best investment grade
-(on a scale of AAA, AA, A, BBB \ldots C).\textsuperscript{$\star$}
+(on a scale of AAA, AA, A, BBB \ldots C).
 
 Returning to our original example, SW18 4HU is a promising choice of
-area for property investment. It's median-priced for the sector, which
-is the cheapest of SW18 (in turn part of the desirable SW London area;
-Figure 1). It has good growth projections of over £28,000 in the next
-year (Figure 2), and receives the top AAA investment grade, being
-in the 90\textsuperscript{th} percentile for predicted price growth
-and low historical volatility in the dataset.
+area for property investment.\textsuperscript{$\star$} It's
+median-priced for the sector, which is the cheapest of SW18 (in turn
+part of the desirable SW London area; Figure 1). It has good growth
+projections of over £28,000 in the next year (Figure 2), and receives
+the top AAA investment grade for being in the 90\textsuperscript{th}
+percentile for predicted price growth and low historical volatility
+compared to all other sectors covered by the dataset.
 
 {\let\thefootnote\relax\footnotetext{\textsuperscript{$\star$}
-    Data intended to assist investors and does not constitute
-    investment advice; independent advice should be sought where appropriate. }}
+    Data intended to assist investors but does not constitute
+    investment advice; independent investment advice should be sought where
+    appropriate. The value of investments can go down as well as up. }}
 
@@
 There are currently no mainsteam property sales or rental sites that
-attempts to provide any quantitative insights to contextualise a
+attempt to provide any quantitative insights that contextualise a
 property's list price. Some list recent nearby sales in plain text,
@@
 richer displays (as demonstrated with the fractal context view,
-\emph{above}), which offer valuable and practical insights to
-prospective homebuyers.
+Section A: Figure 1, and by the interactive data-driven map shown in
+Figure 5), which offer valuable and practical insights to prospective
+homebuyers.
 
@@
 site could be fed-back to help further enhancement and expansion of
-the data-driven insights and visualisations we offer.
+the data-driven insights and visualisations we offer, such as how
+often a given area is searched for on their site and by which
+demographics.
 
@@
 
-\logo \hspace{.1em} would release both a {\bf web service} and associated
-{\bf mobiles and tablet applications}. These would offer limited analysis and
-visualisation without any fee, with collected user data and feedback helping to
-validate the business model and develop our software.
+\logo \hspace{.1em} will release an initial product both as a {\bf web
+  service} and associated {\bf mobiles and tablet applications}
+(Figure 6). These
+offer limited analysis and visualisation without any fee, with
+collected user data and feedback helping to validate the business
+model and futher develop our software.
 
-\begin{wrapfigure}[12]{r}{.4\textwidth}
+\begin{wrapfigure}[14]{r}{.46\textwidth}
+\vspace{.5em}
 \centering
-\includegraphics[width=.39\textwidth]{Figures/mockup.png}
+\includegraphics[width=.42\textwidth]{Figures/mockup.png}
 \caption{ iOS app mock-up.}
@@
 estate investment. Amateur property investors and prospective
-landlords can make data-driven decisions, the likes of which are likely
-already employed by large investment trusts and funds.
+landlords can make data-driven decisions, the likes of which are 
+already employed by analysts for large investment funds.
 
@@
 offered, integrating novel public and privately-acquired datasets and
-release them under a rolling subscription model, aimed at private
-landlords and property speculators, as well as professional real
-estate investment trusts (REITs). Significant revenue could be
-generated through tailored partnerships with funds and high net-worth
-individuals.
+release them to a premium app under a rolling subscription model,
+aimed at private landlords and property speculators as well as
+professional real estate investment trusts (REITs). Significant
+revenue could be generated through tailored partnerships with funds
+and high net-worth individuals.
 
@@
 financial services. Our domain expertise and existing partnerships and
-products would place us as a market leader in this undeveloped
+products would place us as a market leader in this largely undeveloped
 segment.
 
-\section*{Summary}
-
-\begin{wrapfigure}[10]{r}{.3\textwidth}
-\vspace{2em}
+\begin{wrapfigure}[10]{l}{.33\textwidth}
+\vspace{5em}
 \centering
@@
 %\caption*{ {\sf Democratising real estate investment}}
-\caption{ Logo design. }
+\caption{ Branding. }
 %\vspace{-4em}
@@
 
+\section*{Summary}
+
 In this report I've given a brief overview my attempts to elute
-maximal insight from a single large dataset through visualisations and
-quantitative modelling. Together these form a cohesive narrative of
-investigative analysis into a single exemplar postcode, and discovery
-of the potential investment returns therein.
+maximal insight from a single large dataset through statistical
+modelling, visualisations and other data science techniques. Together
+these form a narrative of investigative analysis into a single
+exemplar postcode, and the assessment of the potential investment
+returns within.
 
-These analyses have the potential to generate valuable insights for a range of
-customer segments:
+I have shown that these analyses have the potential to generate
+valuable insights for a range of customer segments:
+
+\setitemize[0]{leftmargin=4em,itemindent=2em}
 \begin{itemize}
@@
 
-\vspace{1em}
-\begin{flushright} {\small An enhanced online version of this short
+\vspace{2em}
+
+\hrule
+
+\vspace{1.5em}
+%\begin{flushright}
+\centering 
+{\small An enhanced online version of this short
     report is available at:
     \leavevmode\href{http://blm.io/datarea}{blm.io/datarea}}
-\end{flushright}
+%\end{flushright}
 

### 2014-11-07 R/arima_model.R

@@ -75,7 +75,7 @@ afit <- data.frame(Month=sdf$Month, median=as.vector(fitted(aa)))
 
 pdf$type <- factor(pdf$type, levels=c("measured", "forecast"))
 
-pdf("plots/forecast_v2.pdf", 5, 3)
+pdf("plots/forecast.pdf", 5, 3)
 ggplot(pdf, aes(x=Month, y=median/1e3)) +
   # forecase bounds
   geom_ribbon(data=vars, inherit.aes=F,

### 2014-11-07 report/sdc_report.tex
final report tweaks

@@
 %	ABSTRACT
-%----------------------------------------------------------------------------------------
+% ----------------------------------------------------------------------------------------
 \vspace{-4em}
@@
 ways. Firstly, the \emph{fractal context} of an area visualises how house
-prices relate to those of neighbouring areas at different scales up
+prices relate to those of neighbouring regions at different scales up
 and down the postcode hierarchy. Secondly, I fit \emph{ARIMA models}
@@
 
-Letting and sales sites sometimes list some recent sales in a given
-area, but it's not currently possible to get a quantitative overview
-of property prices in a region, within the context of its sector,
-district and postcode.
+Property letting and sales sites sometimes list a few recent sales for
+a given area, but it's not currently possible to get an overview
+of the price distribution within an area, or to frame local prices
+within the context of other nearby sectors, district and postcode.
 
@@
 
-These questions can be answered with a clear and intuitive data
+This can be addressed with a clear and intuitive data
 visualisation of price distributions in the postcode hierarchy --- what
-I've called the fractal context of a property price (Figure 1).
+I've called the fractal context of a postcode (Figure 1).
 
 Any chosen area can be seen in the context of its surroundings, from
-full postcodes to much wider areas, clearly showing how the pricing of
+exact full postcodes to much wider areas, clearly showing how the pricing of
 any given property relates to those nearby.  The example graphic for
 postcode SW18 4HU (Figure 1) combines ranked violin density plots and
-a geo-heatmap for a clear, intuitive view of a location's price
+a geo-heatmap for a varied and informative view of a location's price
 context.
@@
 Property investors need a method of evaluating the potential investment
-returns of a given property area. This can be achieved through time
+returns for a given property area. This can be achieved through time
 series modelling with autoregressive integrated moving average (ARIMA)
@@
 which allowed a $drift$ term to capture non-stationary trends
-(i.e. allowing for gain, loss or neither) as well as coefficients
-reflecting periodicity or seasonal effects. Hence potential investment
-returns could then be optimised by selecting areas with
+(i.e. allowing for price gains, losses or neither) as well as coefficients
+reflecting periodicity or seasonal effects. Predicted investment
+returns can then be optimised by selecting areas with
 maximal growth projections.
@@
 \begin{figure}[h]
-\begin{center}
+\centering
+\vspace{.5em}
 \includegraphics[width=.9\textwidth]{Figures/arima.png}
-\caption{Example ARIMA forecast of median house prices for a specific postcode.}
-\end{center}
+\vspace{-.5em}
+\caption{ ARIMA model and forecast of median monthly house prices for the
+  SW18 4 sector.}
+\vspace{.75em}
 \end{figure}
 
-\vspace{-2em}
+\vspace{-1.8em}
 \subsection*{Investment grading}
@@
 variability allows an approximate investment grading of each sector,
-relative to other postcodes within the area covered by this
-dataset. This intuitive output metric can help prioritise areas ripe
+relative to all others covered by this dataset (over 2,500
+sectors). This intuitive output metric can help prioritise areas ripe
 for property investment and combines the theoretical model projections
-with an empirical metric of past stability.
+with an empirical measure of past stability.
 
@@
 
-There are currently no mainsteam property sales or rental sites that
+There are currently no mainstream property sales or rental sites that
 attempt to provide any quantitative insights that contextualise a
@@
 collected user data and feedback helping to validate the business
-model and futher develop our software.
+model and further develop our software.
 
@@
 estate investment. Amateur property investors and prospective
-landlords can make data-driven decisions, the likes of which are 
-already employed by analysts for large investment funds.
+landlords can use our products to make data-driven decisions, the
+likes of which are already employed by analysts for large investment
+funds.
 
@@
 offered, integrating novel public and privately-acquired datasets and
-release them to a premium app under a rolling subscription model,
-aimed at private landlords and property speculators as well as
-professional real estate investment trusts (REITs). Significant
-revenue could be generated through tailored partnerships with funds
-and high net-worth individuals.
+releasing them to a premium app offered under a subscription model,
+aimed at private landlords, professional property developers and
+property speculators as well as real estate investment trusts
+(REITs). Significant revenue could be generated through tailored
+partnerships with funds and high net-worth individuals, likely
+developed from leads that use this service.
 
@@
 In this report I've given a brief overview my attempts to elute
-maximal insight from a single large dataset through statistical
+maximal insights from a single large dataset through statistical
 modelling, visualisations and other data science techniques. Together
 these form a narrative of investigative analysis into a single
-exemplar postcode, and the assessment of the potential investment
-returns within.
+exemplar postcode, culminating in the assessment of potential investment
+returns for properties in this area.
 
